### 2018-12-25 ongoing/report/article.tex
Create project: DS and OMP comparison

\documentclass[12pt]{article}

\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[linkcolor=blue]{hyperref}
\usepackage{xcolor}
\usepackage[ocgcolorlinks]{ocgx2}

\author{Tzu-Yu Jeng}
\date{\today}
\title{Compressive Channel Sensing of Multipath Millimeter Channel via a Beamspace Dantzig Selector and its Probabilitic Performance Bound}

\newcounter{numResult}
%\setcounter{numResult}{1}
\newcommand{\myCount}
{
   \stepcounter{numResult}
   \textbf{\arabic{numResult}}
}

\begin{document}

\maketitle

\section{Acknowledgement}

\section{Abstract}

Multiple-input-multiple-output (MIMO) communication is adopted as the next generation of wireless communication scheme.
However, reliable communication in such r\`egime usually requires that the channel response be known at the receiver, necessary for example in beamforming algorithm and channel calibrupation.
It is then a non-trivial problem to estimate real-world wireless channels

Meanwhile, in the millimeter wave r\`egime, which is usually used in tandem with MIMO, channel often exhibits sparse properties.
If the sparsity is exploited, few observations may suffice to estimate the channel, and such algorithm falls under the study of compressive sensing.
A promiment estimator is The Dantzig Selector proposed by Tao and Cand\'es.
In this article, we extend their proof to derive a specific bound in the setting of estimation of millimeter wave channel, utilizing an all-phase-shifter combiner in the receiver end.
Simulation compares our result and the

\section{Introduction}

\subsection{Background}

Millimeter waves communication has been proposed to be the next-generation specification.
The smaller wavelength entails higher frequency bands, and thus wider available bandwith to be available.
In addition, the smaller closer spaced antennae makes it possible to increase the number antennae.
Multiple-input multiple-output (MIMO) systems with a large number of antennae of both sides (hereafter massive MIMO) is therefore expected to improve spectral efficiency.

But MIMO presents new obstacles as well.
The hardware overhead due to large number of antennas increases complexity and power consumption, and new precoders are being invented to take account of this.
It is important to devise feasible simulation method, as the systems themselves are growing more complicated, and the cost of simulations is large.

The figure of merit of a MIMO system is usually considered to be the channel capacity.
The capacity \(C\) of a MIMO channel \(H\) at perfect CSI (here, the matrix form of \(H\)) is well-known, which is first proved by Teletar.
But that knowledge is hardly always available.
What makes things more difficult is the case that \(H\) is determined from not only \(n(\omega)\), but also parameter \(\theta\).
It is pointed out that the explicit form of \(C(\theta)\) is a difficult and long-standing problem (Goldsmith et.\ al.\ 2003).
Consequently, the perfect-CSI expression of channel capacity is often used regardless of these issues.
Indeed, analysis of precoding algorithm, for example, usually makes use of the sum rate \(\log \det (I +\mathrm{SNR})\).
Therefore the real-time estimation, \(\hat{H}\), of \(H\) is of paramount importance.
Needless to say, when the \(\hat{H}\) is imprecise, resulting analysis is also undermined.

In real applications, channel may also be fast varying with respect to time, and a high complexity algorithm is surely less than being ideal.
With more antennae present, conventional algorithms also increase in complexity and storage.
A design of new algorithms that addressed these issues is thus in need.

\subsection{The Dantzig Selector}

It is a recent developemnt that the estimation of channels is facilitated by advances of compressed sensing.
A series of paper by Cand\`es and Tao (2006) marks its advent.
The idea is that, in many important statsitical applications, the number of variables or parameters \(p\) is much larger than the number of observations \(n\).
In such case of insufficient observations, possibly even with noise, do we have the knowledge of all \(p\) variables?
Of course, more assumption must be made to make the question meaningful.
Cand\`es and Tao's pioneering work reveals the phenomenon that few observations of the signal in question may be sufficient for us to resonstruct the signal when it is sparse in a certain sense.
They argue that such recovery of signals is possible, if we make a few carefully constructed, and seemingly random measurements.
Ever since, it becomes feasible that a camera equipped with few sensors may obtain high quality images, greatly reducing the subsequent cost.

Particularly relevant and inspiring to the present article is Cand\`es and Tao (2006), where they proposed a convex program called The Dantzig Selector (hencefore DS).
Cand\`es and Tao argues in their work that DS has many advantages, and the error probability is upper bounded quantatively and shown to be vanishingly small.
From the realistic perspective, DS formulates the sparse sensing problem as an \(\ell_2\) minimization problem, which is convex, thus techniques from convex optimization may readily be used.
Indeed, they have put the code on the web for the reader to access and verify (Cand\`es \& Romberg 2005).

Several notions related to compressed sensing are necessary in the presentation of results:

\myCount
\textbf{Definition}.
We call \(h \in \mathbb{R}^p\) to be \(S\)-sparse, for \(S =1,2,3,\dotsc\), if only \(S\) components of \(h\) is nonzero.

Denote as \(X_T\) the columns of \(X\) having indices in \(T\), where \(T \subset \{1,\dotsc,p\}\).

\myCount
\textbf{Definition}.
Let \(X\) be an \(n \times p\) matrix having unit \(\ell_2\)-norm columns.
For each integer \(S \in \mathbb{N}\), we say that \(X\) satisfies the RIP of order \(S\) with respect to parameter \(0 \leq \delta_S \leq 1\), if, for all \(\theta\) with \(\|\theta\|_0 \leq S\), for all \(T\) with \(|T| \leq S\),
\begin{gather}
(1-\delta_S) \|\theta\|^2
\leq \|X_T \theta\|_2^2
\leq (1+\delta_S) \|\theta\|^2
\end{gather}
RIP of order \(S\) is essentially saying that the matrix \(X\) is ``almost isometry'' up to ``relative error'' \(\theta_S\).

Moreover, for \(T,T' \subset \{1, \dotsc, p \}\), let \(\theta_{S,S'}\) be the smallest constant such that
\begin{gather}
| \langle X_T c, X_{T'} c' \rangle |
\leq \theta_{S,S'} \cdot \| c \|_2 \|c'\|_2
\end{gather}

It is shown that (Cand\'es and Tao 2005), in the noiseless case, the convex program
\begin{gather}
\min_{\hat{h} \in \mathbb{R}^p}  \|\hat{h}\|_1 \\ \notag
\mathrm{subject}\; \mathrm{to} \quad X \hat{h} =y
\end{gather}
recovers \(\hat{h}\) completely if \(\delta_{S} +\delta_{2S} +\delta_{3S} <1\).

For concreteness, consider a channel with noise,
\begin{gather}
y =\sqrt{\mathcal{E}} X h +z
\end{gather}
where \(z\) is an i.i.d., zero-mean, \(\sigma\)-variance AWGN vector.
(They also discussed specially the noiseless case, but we will not pursue the bounds here.)

How can we hope to estimate \(h\) when, in addition to insufficient observations, there are too few observations?
Let the following linear program, DS, be defined
\begin{gather}
\min_{\hat{h} \in \mathbb{R}^p}  \|\hat{h}\|_1 \\ \notag
\mathrm{subject}\; \mathrm{to} \quad \|X^\dagger r\|_\infty \leq \lambda_p \sigma
\end{gather}
where \(r =y -X \hat{h}\).

DS may be written as
\begin{gather}
\min_{x \in \mathbb{R}^p}  \|x\|_1 \\ \notag
\mathrm{subject}\; \mathrm{to} \quad \|Ax+b\|_\infty \leq c
\end{gather}
And it can be shown that this \(\ell_\infty\)-constraint problem may be recast as a linear program.

The proof of DS is non-constructive, and it is not clear at first what matrices serves as the RIP condition.
But later, () have found there is a particularly convenient sufficient condition to verify RIP.
Tao explicitly remarked that the bound can be refined, but they did not undergo such task.

\subsection{Compressed Channel Sensing}

Meanwhile, physical evidences suggest that millimeter wave channels can be said to be sparse in the frequency domain in a certain senses.
For example, Bajwa et.\ al.\ (2010) argues the \(\ell_0\)-norm of the channel matrix may be bounded by a constant, and in such settings the Dantzig Selector may be applied.
They explore the estimation of single-antenna channel response with respect to time.
Another paper by the same group of scholars (Bajwa et.\ al.\ 2008) shows that \(X\) is RIP for overwhelming probability.

To explain the idea, consider a simplified scenario.
Let \(x[t] \in \mathbb{R}^N\) be random instances of i.i.d.\ Rademacher distribution.
Consider a linear time-invariant channel with
\begin{gather}
y[t] =(x[t] \star h[t]) +z[t]
\end{gather}
where \(h[t] \in \mathbb{R}^L\) is the channel's impulse response, \(x[t] \in \mathbb{R}^N\) as the input, \(y[t] \in \mathbb{R}^N\) as the output, and \(z[t] \in \mathbb{R}^L\) are random instances of i.i.d.\ normalized AWGN, and \(N \gg L\).
Here, if the convolutional relation is expressed by a matrix \(X\), \(X\) is a Toeplitz matrix, so that
\begin{gather}
y =X h
\end{gather}
where \(X\) takes the form
\begin{align}
X
=\begin{bmatrix}
x_1, &0, &\cdots, &0 \\
x_2, &x_1, &\cdots, &0 \\
\vdots, &\vdots, &\ddots, &0 \\
x_N, &x_{N-1}, &\cdots, &x_1
\end{bmatrix}
\end{align}
We seek to get \(h\) in this undertermined system.

Then \(X\) is RIP of overwhelmingly probability.

The literature on CSS has since been vast, and we need not and cannot go through all of them here.
For example, Rao and Lau (2014) considers a rather complicated distributed algorithm where CSS is applied.
They consider a MU-MIMO problem that channels seen by various users share the same support, and proposes a modified joint OMP algorithm.
However, the sparsity of the channel is likewise merely made as an assumption.

Since the notice of hybrid beamforming, scholars have applied

J Lee, Gil, and Lee (2016), as an example, considers a hybrid beamforming system, where the composition of the precoders and combineres play the role of sensing matrix.
Here, as usual, the beambook phase is quantized, and they have specifed a particular set of angles and argues the superiority.

Alkhateeb, Leus, and Heath Jr.\ poses the interesting trade-off of number of OMP measurement and accuracy in an application of OMP.

They work simplified combiner model, where every receiver antenna may only multiply the signal by a constant, and the channel model is assumed to take value on a quantized, non-uniform set of angles.
However, their work mainly focuses on numerical simulation, rather than mathematical proof.

This being said, previous work only makes the assumption on the bound of the norm of the channel matrix, and, to the best knowledge of the author, has not characterized the such probability bounds as involving, say, number of channel paths and vector of array response.

\subsection{Contribution}

This article suggests a convex program analogous of DS in order to directly estimate the channel under hypothesis of uniform linear array response.
Moreover, inspired by the lines of proof of Cand\`es and Tao (2006), the author gives such an asymptotic bound for the probability will be shown, that indicates our method is successful for overwhelming probability.

It is interesting to note that, ever since, the direct application of \(\ell_1\)-problem following the orignal Cand\`es-and-Tao approach has been seemingly rare.

Meanwhile, 
This serves partly as the foundation of performance analysis of 
An analysis of comparison of OMP with DS
Particularly worthy to note

But, in view of the requirement that the sensing matrix should be elementwise i.i.d.\ Gaussian, Tropp and Gilvert (2007) seems to be more restrictive than

Many papers have since cite Tropp and Gilbert (2007) only for justifying bound
But not always is a elementwise i.i.d.\ Gaussian.

After its advent, OMP becomes one of the most widely used channel estimation technique.
The sparsity of channel 
The matrix consisting of the collection of possible response according to available paths --- called the dictionary --- plays the role of the sensing matrix of OMP.
The combiner at the receiver end corresponds to the step where we map the data from high to low dimension.

Alkhateeb et.\ al.\ (2014) suggests an adaptive OMP algorithm.
Given a (beam) code book that corresponds to quantized AoA and AoD, they compare the power of (beam) code words in it, and choosing the column with 
Interestingly, their proof for probability bound of mistaken event does not base on Gilbert and Tropp (2007), but only involves the union bound and direct probabilistic integration.

Meanwhile Cai, Wang, and Xu (2010), making assumptions on MIP matrices, give new bound on OMP, and, in addition, Cai and Wang (2011) replicates the same for analysis for DS, among other convex algorithms.
characterizes on MIP matrices.
Ben-Haim et.\ al.\ follow up their work and refine their bounds, also considering MIP conditions.
The \(\mu\) is easier to verify.
They serve as the guarantee for Alkhateeb et.\ al.\ (2015).

For example, Alkhateeb, Leus, and Heath (2015), being a recent paper, cites Ben-Haim et.\ al.\ as the main foundation for OMP perfomance bound.
Ben-Haim et.\ al.\ conclude that OMP is better when

These works are very technical, and the author does not wish to dispute their accuracy.
Still, it remains unclear which of l1 and greedy is better.
The comparison on Ben-Haim splits the category of convex optimization and greedy.

The author argues that this article nevertheless considers a new configuration that, to the best of our knowledge, has not been dealt with.
First, we do not assume, like Lee et.\ al., that the virtual channel model is naturally quantized in its angle of departure and of arrival.
Accordingly, the effect of quantization of the virtual channel's phase angle incorporated in the beambook has been especially considered in the bound.
The combiner made of relatively phase shifters serves readily as the sensing matrix that satisfy the RIP consideration.
Moreover, an explicit bound has been shown to provide realistic guide for engineering, where the sparsity of the virtual channel matrix, rather than just assumed, is written out as a function of the number of paths of the channel.
Our simple but typical model is expect to shed light on how CSS should be designed in face of phase quantizing and channel parameters available such as path numbers.

Gao et.\ al.\ (2015) discusses the jointly reconstruction of several high-dimensional sparse signals having the same support, using different measurement matrices.
in a modified basis pursuit problem.

Tropp and Gilbert put their techical proof on a standalone online document.
They restrict to i.i.d.\ Gaussian matrix.

\section{Problem Setting}

\subsection{Notation}

Before we start, let us remind that in this article, to avoid confusion all results other than the major theorems to be proved are called lemmata, so the title may be different from the source, but such case will be specified.

We restrict our discussion to the Hilbert space over the field of \(\mathbb{C}\), equipped with usual inner product.
Here the complex numbers has natural interpretation as the phasor of electronic waves.
Boldface lower case Latin alphabets denote vectors, and boldface upper case Latin alphabets denote matrices.

Denote \(\mathbb{E}\)
Since our discussion involving the entire communication system will get very complicated, we shall add subscripts on probabilistic integrals, including \(\mathbb{P}\) and \(\mathbb{E}\).
Let the direct product
%
\begin{gather}
\Omega
=\Omega_t \times \Omega_h \times \Omega_n \times \Omega_r
\end{gather}
%
denote the overall sample space, where:
%
\begin{itemize}
\item \(\Omega_t\) denotes the collection of probability samples at the transmitter end,
\item \(\Omega_h\) denotes that of the communication channel,
\item \(\Omega_h\) denotes that of the noise arising in the communication channel,
\item \(\Omega_r\) denotes that at the receiver end.
\end{itemize}
%
\(\omega_1\) stands for the probability index of the transmitter end, and \(\omega_2\) for the channel, and \(\omega_3\) for the receiver.

\subsection{Channel Model}

We restrict our consideration to the uniform linear array, which can be modeled as
\begin{gather}
\mathbf{a}(\phi')
=\sum_{n=1}^{N} e^{n \phi' i} \mathbf{u}_n
\end{gather}
where \(i\) is the imaginary unit, \(\mathbf{u}_n\) is the unit vector of the \(n\)-th component.

And consider the virtual representation of the MIMO channel (see for example Akdeniz et.\ al.\ 2014).
\begin{gather}
H
=\sum_{l=0}^L \alpha_l \mathbf{a}(\phi_l) \mathbf{a}(\theta_l)^\dagger
\end{gather}
The physical meaning of \(\theta_l\) is the \(l\)-th angle of arrival, and \(\theta_l\) the \(l\)-th angle of arrival.
It is noted that in real scenario, the exponent \(i n \phi_l\) should be multiplied by a factor \((1/\lambda) \sin \eta\), \(\eta\) being the incident angle, accounting for the spacing of two adjacent antenna.
But, since we consider uniform distribution of \(\theta_l, \phi_l\)'s, it suffices to absorb such factors into the values of \(\theta_l, \phi_l\)'s.

Moreover, let the convention for discrete Fourier transform be adopted as below for completeness of presentation.

\myCount
\textbf{Definition.}
Define \(\mathrm{vec}(A) \in \mathbb{R}^{nm}\) to be the vectorization of \(A \in \mathbb{M}(n,m)\).

\myCount
\textbf{Definition.}
For \(\mathrm{vec}(A) \in \mathbb{M}(n,m)\) and \(\mathrm{vec}(A) \in \mathbb{M}(n,m)\)
Define the Kronecker product \(\mathrm{vec}(A) \in \mathbb{R}^{nm}\) 
\myCount
\textbf{Lemma.}

\myCount
\textbf{Definition.} Let \(\mathbf{v} \in \mathbb{R}^m\) be given.
We say \(u \in \mathbb{R}^m\) is the discrete Fourier transform of \(\mathbf{v}\), if
\begin{gather}
\mathbf{u} =F \mathbf{v}
\end{gather}
where each entry of the matrix \(F\) is defined to be
\begin{gather}
F_{k,n} =e^{2\pi ikn/N},
\quad k,n =0,1,2,\dotsc, N-1
\end{gather}

\(H\) has \(N\) columns, but to estimate it, we may agree to send one pilot signal at one time, which consists of only one antenna sending signal, namely \(\mathbf{u}_n\).
This way, we get
\begin{gather}
\tilde{H}_n
:=H \mathbf{u}_n
=\sum_{l=0}^L \alpha_l \exp(-i n \theta_l) \mathbf{a}(\phi_l)
\end{gather}
If we absorb angle of incidence (formed by the ray of electronic wave and the normal line)
\((d/\lambda) \sin \tilde{\theta}_l\)

\subsection{Proposed Algorithm}

\section{Results and Proofs}

\stepcounter{numResult}
\arabic{numResult}
\textbf{Lemma.} Let \(\phi(\omega_3)\) be a uniformly distributed in \([0,2\pi)\) according to probability space index \(\omega_3 \in \Omega\).
And the function \(\mathbf{a}(\phi)\) is defined as in \eqref{}.
Then the discrete Fourier transform

To find the \(l\)-th path's contribution to the frequency domain, recall the identity
\begin{align}
S(\phi)
:=&\left| \sum_{n=0}^{N-1} e^{i n \phi} \right| \\ \notag
=&\left| \frac{e^{i N \phi} -1}{e^{i \phi} -1} \right| \\ \notag
=&\left| \frac{e^{i N \phi/2} -e^{-i N \phi/2}}{e^{i \phi /2} -e^{-i \phi /2}} \cdot e^{i (N-1) \phi/2} \right| \\ \notag
=&\frac{|\sin(N \phi/2)|}{|\sin(\phi /2)|},
\quad 0 \leq \phi \leq \pi
\end{align}
By taking the first two term of the expansion \(\sin x =x -x^3/6 +- \dotsc\), and noting the monotony, 
\begin{align}
S(\phi)
\leq& \frac{1}{\phi/2 -(\phi/2)^3/6} \\ \notag
=& \frac{-48}{\phi^3 -24\phi}
\quad 0 \leq \phi \leq \pi,
\end{align}

The fraction above can be bounded.

By the same token, more generally,
\begin{align}
&\left| \frac{1}{N} \sum_{n=0}^{N-1} e^{-2\pi i k n/N} e^{i n \phi} \right| \\ \notag
=&\frac{|\sin( \phi N /2 )|}{N |\sin( \pi k /N -\phi /2 )|} \\ \notag
=&\frac{1}{N} S(\pi \phi') \\ \notag
\leq& \frac{48}{N \pi^3 (\phi'^3 -24 \phi' /\pi^2)}
\end{align}
where \(\phi' =k/N -\phi/2\pi\), and the cubic expression is split.
We wish to bound the tail of \(S(\pi \phi')\), and apply the na\"ive bound
\begin{align}
&\frac{N}{2\pi} \cdot \frac{48}{N \pi^3} \int_{L/N}^1 \frac{1}{\phi'^3 -24 \phi' /\pi^2}
 d\phi' \\ \notag
=&\frac{1}{2\pi^2} \cdot \left( 2 \log \frac{L}{N} +\log \frac{24-\pi^2}{24-\pi^2 L^2 /N^2} \right) \\ \notag
\leq& \frac{1}{\pi^2} \log \frac{L}{N}
\end{align}
Here, the integral in the first line is immediate.
The second term in the bracket is smaller than 0 and is dropped in the last inequality.
The bound is satisfying enough, though refinement is surely possible.
Indeed, \(\log L/N \to 0\) as \(N \to \infty\) and limited by the number of paths \(L\).

%A =\mathrm{supp} \beta

where
\begin{gather}
\mathcal{A}, \mathcal{B}, \mathcal{C} \subset \mathbb{N} \\ \notag
\mathcal{A} \cap \mathcal{B} =B \cap C =A \cap C =\varnothing \\ \notag
A \cup B \cup C =\{1, \dotsc, N\}.
\end{gather}

\textbf{Proposition}.
Let \(h \in \mathbb{R}^n\) and \(\eta \in \mathbb{R}^p\) are defined as in \eqref{}.
Then
\begin{align}
\|h_A\|_1 -\|\eta_A\|_1 +\|\eta_C\|_1 -\|h_C\|_1
\leq& \|h +\eta\|_1 \\ \notag
\leq& \|h\|_1
\end{align}

\textbf{Proof}.

\begin{align}
\|X^\dagger_{AB} X \eta\|_\infty
<&\|X^\dagger X \eta\|_\infty \\ \notag
\leq& 2 \lambda_p
\end{align}

From ``Dantzig Selector'' (Cand\`es and T Tao 2007), Lemma 1, first equation.

\stepcounter{numResult}
\arabic{numResult}
\textbf{Lemma}.
\begin{align}
\|\eta_Ah\|_2
\leq K_1 \|X^\dagger_{AB} X \eta\|_2 +K_2 \|\eta_C\|_2
\end{align}

From ``Dantzig Selector'' (Cand\`es and T Tao 2007), Lemma 1, second equation.

\stepcounter{numResult}
\arabic{numResult}
\textbf{Lemma}.
\begin{align}
\|\eta\|_2^2
\leq \|\eta_{AB}\|_2^2 +K_3 \|\eta_C\|_2^2
\end{align}

Another difference is that this article solves the optimization problem solely using the method of greatest descent.
While OMP is widely used 
With the analysis of [], OMP is usually used

\section{Simulation}

Since our setting is slightly different
In this section we first describe the DS algorithm, which is our target of investigation, and briefly review the OMP algorithm, for sake of comparison.
As we see in the above, Dantzig Selector 
We transform the \(\ell_1\)-minimization problem to a primal-dual algorithm, which is well-known for convex optimization problems.

% % % % % % % % % % % % % % % %

\section{Conclusion and Future Work}

Firstly,
Meanwhile, DS goes not without criticism for large complexity.
Indeed it is clear, even from the primal-dual implementation above, that DS requires large dimensional vector

Secondly
And comparison of bounds of OMP and DS
The author criticises the tendency of current literature to mix up results for DS and OMP without clear justification, since they are two quite different approaches to the compressed sensing problem.
It is however not trivial to make meaningful comparison, since DS works with a RIP matrix, and OMP for Gaussian... the distinction is slight but definite...
The author critizises In OMP we don't know whether the measurement matrix is the same as channel
And OMP has stricter condition than DS
Tao, however, specifies a noise relation

Bounds of performance in different setting are often nontrivial to obtain, and one cannot think OMP, or indeed every algorithm, as a universal solution.
OMP trades off precision for time complexity, and for DS the other way around is true, hence we cannot say which is definitely better than the other.
All in all, different problems requires different technique and algorihm within the constraint at hand, as is probably true in every discipline of engineering.

\section{References}

References are sorted first with alphabetical order, then, in case for exactly the same author(s), with year.

\begin{enumerate}

\item M R Akdeniz et.\ al.\ (2014), ``Millimeter Wave Channel Modeling and Cellular Capacity Evaluation'', \textit{IEEE Journal on Selected Areas in Communications}, Vol.\ 32, No.\ 6

\item A Alkhateeb, O E Ayach, G Leus, and R W Heath Jr (2014), ``Channel Estimation and Hybrid Precoding for Millimeter Wave Cellular Systems'', \textit{IEEE Journal of Selected Topics in Signal Processing}, Vol.\ 8, No. 5.

\item A Alkhateeb, G Leus, and R W Heath Jr.\ (2015), ``Compressed Sensing Based Multi-User Millimeter Wave Systems: How Many Measurements are Needed?'' \textit{IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP)}

\item W U Bajwa, J Haupt, A M Sayeed, and R Nowak (2010), ``Compressed Channel Sensing: A New Approach to Estimating Sparse Multipath Channels'', Proceedings of the IEEE, Vol.98, No.6.

\item W U Bajwa, J Haupt, G Raz, and R Nowak (2008), ``Compressed Channel Sensing'', 2008 42nd Annual Conference on Information Sciences and Systems.

\item R Baraniuk, M Davenport, R DeVore, and M Wakin (2008), ``A Simple Proof of the Restricted Isometry Property for Random Matrices'', \textit{Constructive Approximation} \textbf{28}: 253–263

\item Zvika Ben-Haim, Y C Eldar, and M Elad (2010), ``Coherence-Based Performance Guarantees for Estimating a Sparse Vector Under Random Noise'', \textit{IEEE Transactions on Signal Processing}, Vol.\ 58, No.\ 10.

\item S Boyd, L Vandenberghe (2004), \textit{Convex Optimization}. Cambridge U. Press.

\item E J Cand\`es, J Romberg, and T Tao (2006), ``Robust Uncertainty Principles: Exact Signal Reconstruction From Highly Incomplete Frequency Information'' \textit{IEEE Transactions on Information Theory}, Vol.52, No.2.

\item E J Cand\`es and J Romberg (2005), ``\(\ell_1\)-MAGIC: Recovery of Sparse Signals via Convex Programming'', Retrieved from \url{https://statweb.stanford.edu/~candes/l1magic/downloads/l1magic.pdf}.

\item E J Cand\`es and T Tao (2005), ``Decoding by Linear Programming'', \textit{IEEE Transactions on Information Theory}, Vol.51, No.12.

\item E J Cand\`es and T Tao (2006), ``Near-Optimal Signal Recovery From Random Projections: Universal Encoding Strategies?'' \textit{IEEE Transactions on Information Theory}, Vol.52, No.12.

\item E Cand\`es and T Tao (2007), ``The Dantzig Selector: Statistical Estimation when \(p\) is Much Larger than \(n\)'', \textit{The Annals of Statistics}, Vol.35, No.6.

\item T T Cai and L Wang (2011), ``Orthogonal Matching Pursuit for Sparse Signal Recovery With Noise'', \textit{IEEE Transactions on Information Theory}, Vol.\ 57, No.\ 7.

\item T T Cai, L Wang, and G Xu (2010), ``Stable Recovery of Sparse Signals and an Oracle Inequality'', \textit{IEEE Transactions on Information Theory}, Vol.\ 56, No.\ 7.

\item M P Friedlander and M A Saunders (2007), ``Discussion: The Dantzig Selector: Statistical Estimation when \(p\) is Much Larger than \(n\)'', \textit{The Annals of Statistics}, Vol.\ 35, No.\ 6.

\item Z Gao, L Dai, Z Wang, and S Chen (2015). ``Spatially Common Sparsity Based Adaptive Channel Estimation and Feedback for FDD Massive MIMO'', \textit{IEEE Transactions on Signal Processing}, Vol.\ 63, No.\ 23.

\item A Goldsmith, S A Jafar, N Jindal, and S Vishwanath (2003), ``Capacity Limits of MIMO Channels'', \textit{IEEE Journal on Selected Areas in Communications}, Vol.\ 21, No.\ 5.

\item R W Heath Jr, N González-Prelcic, S Rangan, W Roh, and A M Sayeed (2016), ``An Overview of Signal Processing Techniques for Millimeter Wave MIMO Systems'', \textit{IEEE Journal of Selected Topics in Signal Processing}, Vol.\ 10, No.\ 3.

\item J Lee, G-T Gil, and Y H Lee (2016), ``Channel Estimation via Orthogonal Matching Pursuit for Hybrid MIMO Systems in Millimeter Wave Communications'', \textit{IEEE Transactions on Communications}, Vol.\ 64, No.\ 6.

\item H Li, Q Gao, R Chen, R Tamrakar, S Sun, and W Chen, ``Codebook Design for Massive MIMO Systems in LTE'', \textit{2016 IEEE 83rd Vehicular Technology Conference (VTC Spring)}.

\item X Rao and V K N Lau (2014), ``Distributed Compressive CSIT Estimation and Feedback for FDD Multi-User Massive MIMO Systems'', \textit{IEEE Transactions on Signal Processing}, Vol.\ 62, No.\ 12.

\item J A Tropp and A C Gilbert (2007a), ``Signal Recovery from Random Measurements via Orthogonal Matching Pursuit: The Gaussian Case'', Caltech, ACM Tech.\ Rep. Retrieved from: \url{www.acm.caltech.edu/~jtropp/reports/TG07-Signal-RecoveryTR.pdf}

\item J A Tropp and A C Gilbert (2007b), ``Signal Recovery From Random Measurements via Orthogonal Matching Pursuit'', \textit{IEEE Transactions on Information Theory}, Vol.\ 53, No.\ 12.

\end{enumerate}

\end{document}