### 2019-05-29 dantzig_selector/presentation/5_slide.tex

\Frame {Conclusion}
{
\I We are concerned with effective and estimation of MIMO mm-wave channel by exploiting its sparsity

\I Towards that end, We adopt The Dantzig Selector (DS) rather than the popular Orthogonal Matching Pursuit (OMP)

\I We generalized DS for complex vectors, and carried it out in the angular domain

\I Analogous to original DS proof, we bound the expected square error, and the successful probability is also bounded by a explicit constant
}

\page [yes] % % % % % % % % % % % % % % % % % % % % % % % % % %

\Frame {Future Work}
{
\I Check the proof carefullly, especially approximation

\I Finish the simulation program

\I Compare the probability bound with simulation result

\I Also implement OMP and compare they two
}

\page [yes] % % % % % % % % % % % % % % % % % % % % % % % % % %

\Frame {References}
{
{\tfxx
\I D Achlioptas (2001). Database-friendly random projections. In Proceedings of the twentieth ACM SIGMOD-SIGACT-SIGART symposium on Principles of database systems (pp. 274-281). ACM.

\I W U Bajwa, J Haupt, G Raz, and R Nowak (2008), ``Compressed Channel Sensing'', 2008 42nd Annual Conference on Information Sciences and Systems.

\I R Baraniuk, M Davenport, R DeVore, and M Wakin (2008), ``A Simple Proof of the Restricted Isometry Property for Random Matrices'', \It{Constructive Approximation} \Bf{28}: 253–263

\I E J Cand\`es and T Tao (2005), ``Decoding by Linear Programming'', \It{IEEE Transactions on Information Theory}, Vol.51, No.12.

\I E Cand\`es and T Tao (2007), ``The Dantzig Selector: Statistical Estimation when \m {p} is Much Larger than \m {n}'', \It{The Annals of Statistics}, Vol.35, No.6.
}
}