### 2019-06-05 dantzig_selector/report/3_theorems_and_proofs.tex

\startchapter [title={Theorems and Proofs}]

\input {./3_1_almost_sparsity.tex}
\input {./3_2_confirming_restricted_isometry.tex}
\input {./3_3_expected_error_analysis.tex}

\stopchapter