### 2019-06-09 dantzig_selector/report/3_theorems_and_proofs.tex
Proof: Confirming RIP: Introduce more indication function

\startchapter [title={Theorems and Proofs}]

%\input {./3_1_almost_sparsity.tex}
\input {./3_2_confirming_restricted_isometry.tex}
%\input {./3_3_expected_error_analysis.tex}

\stopchapter