### 2019-08-18 dantzig_selector/presentation_B/5_conclusion.tex

\Frame {Conclusion}
{
\I We are concerned with effective and estimation of MIMO mm-wave channel by exploiting its sparsity, and studied Dantzig Selector (DS).

\I We modified DS for present setting: for almost sparsity, in the angular domain, for the almost-sparse condition, for the beamformer matrix in our setting, and for complex vectors.

\I We bound the expected square error and the successful probability, in explicit expression of system parameters such as \m {N_H}, \m {N_Y}, and \m{L}.
}
% % % % % % % % % % % % % % % % % % % % % % % % % %
\Frame {Future Work}
{
\I Check the proof again, especially the approximation decision.

\I Simulate many different parameters such as \m {N_H}, \m {N_Y}, and \m{L} and discuss extensively.

\I Verify (if correct) the successful-probability bound and observe how \m {N_Y, N_H} affects it (if any).

\I Also implement (some or all of) OMP, LS, LASSO among others, and compare them.
}
% % % % % % % % % % % % % % % % % % % % % % % % % %
\Frame {References}
{
{\tfx
\I D Achlioptas (2001). Database-friendly random projections. In Proceedings of the twentieth ACM SIGMOD-SIGACT-SIGART symposium on Principles of database systems (pp. 274-281). ACM.

\I W U Bajwa, J Haupt, G Raz, and R Nowak (2008), ``Compressed Channel Sensing'', 2008 42nd Annual Conference on Information Sciences and Systems.

\I R Baraniuk, M Davenport, R DeVore, and M Wakin (2008), ``A Simple Proof of the Restricted Isometry Property for Random Matrices'', \It{Constructive Approximation} \Bf{28}: 253–263

\I E J Cand\`es and T Tao (2005), ``Decoding by Linear Programming'', \It{IEEE Transactions on Information Theory}, Vol.51, No.12.

\I E Cand\`es and T Tao (2007), ``The Dantzig Selector: Statistical Estimation when \m {p} is Much Larger than \m {n}'', \It{The Annals of Statistics}, Vol.35, No.6.
}
}