### 2019-09-22 paper/4_simulation.tex

\startchapter [title={Simulation}]

After the introduction of problem configuration and analysis of expected square error, we shall formulate our algorithm and focus on the practical aspects.
As of now, we have transformed the sparse-recovery problem of complex matrix \m {H} to complex vector \m {h}, and finally to real vector \m {\MC {R} \SB {h}}.

It is mentioned in ``\m {\ell_1}-MAGIC'' (Cand\`es and J Romberg 2005) that the complex DS can be converted as an SOCP program.
If so, primal-dual interior point algorithm (for example) may readily be applied on it to simplify the program, for which we may consult Boyd and Vandenberghe (2004), {\it Convex Optimization} chap.\ 11 as a standard reference.
But we will not implement that here.
For brevity, we use Python Library CVXPY to solve the complex DS.

\startsection [title={Parameters}]

We do three series of experiments, which are of size from small to large.
In three experiments, we take respectively \m {L} to be \m {2, 3, 4}, \m {N_Y} to be \m {5, 7, 9}, \m {N_R} to be \m {10, 14, 18}, and \m {N_H} to be \m {20, 28, 36}.

Other parameters are fixed in these experiments; 
The number of grid of quantization of analog precoder is 16
The wavelength of the EM wave that antenna is emitting, \m {\l _{\Rm{ant}} = 1}, is set to be 0.1.
and the antenna spacing \m {d _{\Rm{ant}} =3}.
Note that only their ratio matters.
See chapter 2 for meaning of them.

There are also parameters related to the convex optimization algorithm.
The tolerance of absolute error in the primal-dual algorithm used in CVXPY is set as \m {10^{-4}}, and the tolerance of relative error, as \m {10^{-3}}.
The maximal number of iteration of CVXPY is set to be \m {32}.
On the other hand, the maximal number of iteration of OMP is set to be \m {4 N_H}.

Each data point is repeated for 32 times, and taken average.
We plot the absolute log error
\DispN {
\NC \eta_{\Rm{abs}}
=\NC \RB {\log_2 {\VNm {\V {h} -\Hat {\V {h}}}_2}} _{\Ss{avg}}, \NR
}
and relative log error.
\DispN {
\NC \eta_{\Rm{rel}}
=\NC \F {\RB {\log_2 {\VNm {\V {h} -\Hat {\V {h}}}_2}} _{\Ss{avg}}}
{\RB {\log_2 {\VNm {\V {h}}_2}} _{\Ss{avg}}} \NR
}
The averaged error values are plotted along with the theoretical expected error of DS, derived in chap.\ 3, namely
\DispN {
\NC \eta_{\Rm{th}}
=\NC \F{5}{2} + \log \RB{L N_H} + \log \log N_H \NR
}

Several threshold values occur in relevant alorithms.
The threshold used in DS program is called \m {\g_{\Rm {DS}}} here, and another threshold of Lasso, called \m {\a}.
Yet another threshold used for the residue norm used in OMP program is called \m {\eta}.
Various several values are tried for them, determined as thus: Take a central value roughly corresponding to the theory, and mutiply or divided the central value by powers of 2, or

Concretely, we set \m {\g_{\Rm {DS}} = 2 \R {\log N_H}}, corresponding to Cand\`es \& Tao (2007)
For comparison, \m {\g_{\Rm {Lasso}} = 2 \R {\log N_H}} too.
For OMP, suggested values in Cai \& Wang (2011) is illuminating.
There, Theorem 7 considered \m {\ell _2}-norm, and Theorem 8, \m {\ell _\infty}-norm.
We make simplification and set respectively \m {\R {3} N_Y} and \m {2 \R {\log N_H}}

\stopsection

\startsubsection [title={On Complexity}]

In general, complexity analysis of convex program is difficult, in that it is not easily ensured how many iteration the program uses.
Specifically, the implementers of CVXPY, Diamond and Boyd \cite [DiB16], alleges that the program is usually converted in to primal-dual problem.

However, if we use Newton method and assume self-concordance, there is an analysis in B \& V.
Such analysis, in our opinion, still gives us an idea about the complexity of the DS program.

It can be shown that DS can be cast into an Second Order Cone Programming (SOCP), and it is clear that self concordance applies to SOCP.

The function being minimized is \m {\underset {x} {\Rm{min}} \VNm {\V {g}}_1}.
Let \m {\V {g} _0} denote the starting value of \m {\V {g}}, and \m {\V {g} ^{\star}} the the point of convergence.
Then the number of iteration of Newton method is bounded by
\DispN {
\NC \F {\VNm {\V {g}} _1 - \VNm {\V {g} ^{\star}}} {\g} + \log_2 \log_2 \F {1} {\e} \NR
}
Here \m {\g} is a parameter related to the Backtracking algorithm, and \m {\e} the tolerance of error.
The details can be found in Boyd \& Vandenberghe (2004).
They cites various numerical experiements, and it turns out almost always that a value of \m {\g} close to 1 gives the best result, so we take \m {\g} as a constant.

Numerically, we see that the complexity of DS is generally very high.
But it also depends on the tolerence of error.
% XXX criticism on DS complexity?

\startsection [title={Discussion}]

On the other hand, DS goes not without criticism for its large complexity (Friedlander and Saunders 2007).
Indeed it is clear, even from the primal-dual implementation above, that DS requires memory for large dimensional vector, but OMP does not.

And a meaningful comparison of bounds of OMP and DS is not easy, as their settings are somewhat different.
The author thus criticises the tendency of current literature to mix up results for DS and OMP without clear justification.
Indeed, DS calls for a RIP matrix and guarantees such performance even in noisy observation (Cand\`es and Tao 2005, 2007), while RIP is not easily to justify rigorously, and its easy construction is even an open problem.
On the other hand, as of OMP, besides the original justification of entrywise i.i.d.\ Gaussian sensing matrix (Tropp and Gilbert 2007a), and existent attempt to characterize MIP condition, there is much to be done OMP for Gaussian.

\stopsection

\stopchapter