### 2019-09-29 paper/conclusion.tex

\startchapter [title={Conclusion}]

This article addresses the problem of effective and estimation of the millimeter wave, multiple-input, multiple-output channel.
We adopts the Dantzig Selector (DS) rather than Orthogonal Matching Pursuit (OMP), like many of past literature, as the estimation algorithm.

We proceed the estimation algorithm by exploiting its sparsity in spacial Fourier domain.
We bounded the expected square error, and in doing so, following the spirit of Cand\`es and Tao \cite [CaT07], we characterizes the spacial sparsity, confirmed the restricted isometry of measurement matrix \m {\M {P}}, and justified the generalization for complex vectors space.
The bound assumes a fairly general setting, and is expressed as an explicit function of physical parameters, such as the number of path \m {L} and channel dimension \m {N_H}.

Simulation is done, and DS's performance is compared to OMP, Lasso, and LS.
In our case, DS performs better but gives higher complexity, and our analysis agrees the numerical experiment.
Different problems requires solution within the feasible constraint, as is probably true in every discipline of engineering.
Put concisely, no algorithm is a universal solution, and they trade off precision for time and space complexity, or conversely, among other things.
We hope the investigation of DS can be a welcomed alternative added to the present literature of channel estimation.

\stopchapter