### 2019-10-30 paper/3_3_generalizing_the_dantzig_selector_bound.tex
Finish 3.3

\startsection [title={Generalizing the Dantzig Selector Bound}]
\startsubsection [title={Technical Lemmata in Complex Case}]

The third and last part of our task is to plug in the almost-sparsity found in Section 3.1, and plug in the probablity found in Section 3.2.
Set for short
%
\DispNum {d-d-gg-gg} {
\NC \V {d} 
= \NC \Hat {\V {g}} -\V {g} \NR
}
%
And our task is to bound the expectation \m {\MB {E} \SB {\V {d}}}.

We argue that \m {\V {g}} can essentially be seen as sparse, and for that purpose we ignore relevent terms.
Also, we do not seek a very tight bound and will approximate \m {\d_{S}} by \m {1}.
And, as shall be seen, it suffices to set
%
\DispNum {g'-g'-2N-Nh} {
\NC \g
= \NC \R {4 \log N_H} \NR
}
%
The lemmata below are taken from \cite [CaT07] and modified slightly in accordance to our settings, except possibly the generalization to complex vector.
%
\Result
{Lemma}
{
%
\DispNum {d-1-dA-C1} {
\NC \VNm {\V {d} _{\SB{\MC {C}}}} _1
\leq \NC \VNm {\V {d} _{\SB{\MC {A}}}} _1
+2\VNm {\V {g} _{\SB{\MC {C}}}} _1 \NR
}
}
%
To show this, observe that with triangle inequality applied respectively on \m {\MC {A}} and \m {\MC {C}}, while recalling that \m {\V {g} =\V {g} _{\SB{\MC {A}}}},
%
\Disp {
\NC \VNm {\V {g}} _1
-\VNm {\V {d} _{\SB{\MC {A}}}} _1
+\VNm {\V {d} _{\SB{\MC {C}}}} _1
-\VNm {\V {g} _{\SB{\MC {C}}}} _1
\leq \NC \VNm {\V {g} +\V {d}} _1 \NR
%
\NC =\NC \VNm {\hat {\V {g}}} _1 \NR
}
%
On the other hand, by construction that \m {\hat {g}} minimizes the \m {\ell_1}-norm,
%
\Disp {
\NC \leq \NC \VNm {\V {g}} _1 \NR
}
%
To prove the next result, fix an instance of \m {\M {P}} and instance of \m {\V {z}}.
We claim that
%
\DispNum {E-h-2N-Nh} {
\NC \MB {E} \SB {\Nm {\IP {\V {z} , \M {P} _{\SB {:, n_h}}}}}
\leq \NC  2 \R {2 \log N_H} \NR
%
\NC n_h 
=\NC 1, \ldots, N_H^2. \NR
}
%
This follows by the fact that \m {\M {P} _{\SB {:, n_g}}} has unity \m {\ell_2}-norm, and on the randomness of \m {\V {z}}, the quantity \m {\Nm {\IP {\V {z} , \M {P} _{\SB {:, n_h}}}}} observes Standard Normal.
To bound the failure probability, recall the tail bound for \m {Q} function
%
\DispNum {Q-x-12-22} {
\NC Q\SB {x'}
\leq \NC \F {1} {2} \Ss {e}^{-x'^2 /2} \NR
}
%
Particularly, for \m {x' =\R {4 \log N_H}}, this gives the failure probability
%
\DispNum {Q-h-1N-Nh} {
\NC Q\SB {\R {4 \log N_H}}
=\NC \F {1} {N_H^2}. \NR
}
%
\Result
{Lemma}
{
If \m {\M {P}} is \m {\d_S}-RIP,
%
\DispNum {P-d-4N-Nh} {
\NC \VNm {\M {P}^\Adj \M {P} \V {d}} _\infty
\leq \NC  4 \R {2 \log N_H} \NR
}
%
holds for probability \m {p} 
%
\DispNum {1-p-NH-H2} {
\NC 1 -p
\leq \NC N_H^{-2}. \NR
}
}
%
To show this, by definition
%
\DispNum {z-y-Pg-ny} {
\NC \IP {\V {z} -\RB {\V {y} -\M {P} \hat {\V {g}}}, \M {P} _{\SB {:,n_y}}}
= \NC \IP {\M {P} \hat {\V {g}} -\M {P} \V {g}, \M {P} _{\SB {:,n_y}}} \NR
%
\NC = \NC \IP {\M {P} \V {d}, \M {P} _{\SB {:,n_y}}} \NR
}
%
By construction
%
\DispNum {P-g-Py-Nh} {
\NC \VNm {\M {P} _{\SB {:,n_y}}^\Adj \RB {\V {y} -\M {P} \hat {\V {g}}}} _\infty
\leq \NC \VNm {\M {P}^\Adj \RB {\V {y} -\M {P} \hat {\V {g}}}} _\infty \NR
%
\NC \leq \NC 2\R {2 \log N_H} \NR
}
%
By triangle inequality, together with \Rf {Q-h-1N-Nh}, we obtain the desired result.
%
\DispNum {P-d-zP-Nh} {
\NC \VNm {\M {P} _{\SB {:,n_y}}^\Adj \M {P} \V {d}} _\infty
\leq \NC \VNm {\IP {\V {z}, \M {P} _{\SB {:,n_y}}}} _\infty
+\IP {\V {y} -\M {P} \hat {\V {g}}, \M {P} _{\SB {:,n_y}}} \NR
%
\NC \leq \NC 2 \R {2 \log N_H} +2 \R {2 \log N_H} \NR
}
%
Lastly, the following two lemmata is taken from \cite [CaT07], Lemma 1.
The generalization from real vector spaces to complex spaces is completely straightforward, while meaning of magnitude and inner product has to be changed.
One of them reads
%
\DispNum {d-2-dA-12} {
\NC \VNm {\V {d}} _2^2
\leq \NC \VNm {\V {d} _{\SB {\MC {AB}}}} _2^2 +\F {1} {S} \VNm {\V {d} _{\SB{\MC {C}}}} _1^2 \NR
}
%
By power mean inequality, we get a statement cruder than \Rf {d-2-dA-12} that
%
\Result
{Lemma}
{
%
\DispNum {d-2-dA-C1} {
\NC \VNm {\V {d}} _2
\leq \NC \VNm {\V {d} _{\SB {\MC {AB}}}} _2 +\F {1} {\R {S}} \VNm {\V {d} _{\SB{\MC {C}}}} _1 \NR
}
}
%
The other one of them reads
%
\DispNum {d-2-11-C1} {
\NC \VNm {\V {d} _{\SB{\MC {AB}}}} _2
\leq \NC \F {1} {1-\d_{2S}} \VNm {P _{\SB {\MC {AB}}}^\Tr P d} _2
+\F {\T {\d}_{S, 2S}} {\RB {1-\d_{2S}} \R {S}} \VNm {d_{\SB{\MC {C}}}} _1 \NR
}
%
Here \m {\T {\d} _{s_1, s_2}} is called the \m {s_1,s_2}-restricted isometry constant, and is defined in \cite [Can05] (which uses \m {\th}), as follows.
Let \m {\M {Q}'} be fixed.
For all \m {s_1}-sparse \m {\V {x}_1}, for all \m {s_2}-sparse \m {\V {x}_2}, whose positions for nonzero components are disjoint,
%
\Disp {
\NC \T {\d} _{s_1, s_2}
= \NC \inf \CB {
\d':\; \Nm {\IP {\M {Q}' \V {x}_1, \M {Q}' \V {x}_2}}
   \leq \d' \VNm {\V {x}_1} \VNm {\V {x}_2}
} \NR
}
%
Here we use a result in \cite [Can05] that relates \m {\T {\d} _{s_1, s_2}} to \m {\d_s}, that is
\DispNum {d-2-ds-s2} {
\NC \T {\d} _{s_1, s_2}
\leq \NC \d_{s_1+s_2} \NR
}
And we need another crude bound to relate \m {\d_{2S}} to \m {\d_{S}}.
By triangle inequality, by power-mean inequality,
\DispNum {Q-x-Qx-x2} {
\NC \VNm {Q' \RB {\V {x}_1 + \V {x}_2}}
%
\leq \NC \VNm {Q' \V {x}_1}
+\VNm {Q' \V {x}_2} \NR
%
\NC \leq \NC \RB {1 +\d _{s'}} \RB {\VNm {\V {x}_1} +\VNm {\V {x}_2}} \NR
%
\NC \leq \NC \RB {1 +\d _{s'}} \D \R{2} \VNm {\V {x}_1 +\V {x}_2} \NR
}
Therefore
\Disp {
\NC \d_{2s'}
\leq \NC \R {2} \d_{s'} \NR
}
By the same token,
%
\Result
{Lemma}
{
If \m {\M {Q}'} has \m {\d_{s'}}-RIP, then, for \m {M=1,2,3,\dots} so that \m {\d _{Ms'}} is meaningful,
%
\DispNum {d-s-Md'-d's} {
\NC \d_{Ms'}
\leq \NC \R {M} \d_{s'} \NR
}
}
%
In particular,
%
\DispNum {d'-S-Ll-d's} {
\NC \d_{S}
\leq \NC \R {L} \RB {\log N_H} ^{1/4} \d_{s} \NR
}
%
So \Rf {d-2-11-C1} becomes
%
\Result
{Lemma}
{
If \m {\M {P}} is \m {\d_S}-RIP,
%
\DispNum {d-2-11-C1} {
\NC \VNm {\V {d} _{\SB{\MC {AB}}}} _2
\leq \NC \F {1} {1-\R{2} \d_{S}} \VNm {P _{\SB {\MC {AB}}}^\Tr P d} _2
+\F {\R{3} \d_{S}} {\RB {1-\R{2} \d_{S}} \R {S}} \VNm {d_{\SB{\MC {C}}}} _1 \NR
}
}
\stopsubsection

\startsubsection [title={Main Result}]

We are going to combine previous lemmata and show the main result.
For simplicity, set
%
\DispNum {d'-S-18-18} {
\NC \d_S
\leq \NC \F {1} {8} \NR
}
%
Thus
\DispNum {S-S-4N-NH} {
\NC S
=\NC L \log N_H \NR
}
%
\Result
{Theorem}
{
Let \m {\V {y}}, \m {\M {P}}, \m {\V {g}}, \m {\hat {\V {g}}}, \m {\V {d}} be defined as above.
Then the bound
%
\DispNum {d-2-42-NH} {
\NC \VNm {\V {d}} _2
%
\leq \NC 3.29 \R {L} \log N_H +4.56 \R {L} \R {\log N_H}^3 \NR
}
%
holds for probability \m {p}, with
%
\DispNum {1-p-ON-13} {
\NC 1 -p
=\NC \MC {O} \SB {N_H ^{-1/3}} \NR
}
}

To prove \Rf {d-2-42-NH}, we simplify the expression for \m {\VNm {\V {d} _{\SB{\MC {AB}}}} _2} in \Rf {d-2-11-C1} and the one for \m {\VNm {\V {d} _{\SB{\MC {C}}}} _1} in \Rf {d-1-dA-C1} and 
In this section we substitute rough numbers.

By the definition of truncation, by \m {\ell_p} norm inequality, by \Rf {P-d-4N-Nh},
%
\DispNum {P-2-PP-NH} {
\NC \VNm {\M {P} _{\SB {\MC {AB}}}^\Tr \M {P} \V {d}} _2
\leq \NC \VNm {\M {P}^\Tr \M {P} \V {d}} _2 \NR
%
\NC \leq \NC \R {S} \VNm {\M {P}^\Tr \M {P} \V {d}} _\infty \NR
%
\NC \leq \NC 5.65 \R {\log N_H} \NR
}
%
By \m {\ell_p}-norm inequality, by the definition of truncation, by \Rf {d-2-11-C1}, and by \Rf {P-2-PP-NH} just above,
%
\Disp {
\NC \VNm {\V {d} _{\SB{\MC {A}}}} _1
\leq \NC \R {S} \VNm {\V {d} _{\SB{\MC {A}}}} _2 \NR
%
\NC \leq \NC \R {S} \VNm {\V {d} _{\SB{\MC {AB}}}} _2 \NR
%
\NC \leq \NC 1.21 L \R {\log N_H}^3
+0.262 \VNm {\V {d} _{\SB{\MC {C}}}} _1 \NR
}
%
And \Rf {a-2-4N-NH} gives, after arrangement,
%
\DispNum {d-A1-16-H2} {
\NC \VNm {\V {d} _{\SB{\MC {A}}}} _1
%
\leq \NC 1.64 L \R {\log N_H}^3 +0.951 L \RB {\log N_H}^2 \NR
}
%
Then, substituting \Rf {g-C-4p'-NH} and \Rf {d-A1-16-H2} above into \Rf {d-1-dA-C1}, we get
\DispNum {d-C1-16-H2} {
\NC \VNm {\V {d} _{\SB{\MC {C}}}} _1
%
\leq \NC 1.64 L \R {\log N_H}^3 +3.61 L \RB {\log N_H}^2 \NR
}
%
And \Rf {d-2-11-C1} becomes
%
\DispNum {d-2-16-H3} {
\NC \VNm {\V {d} _{\SB{\MC {AB}}}} _2
%
\leq \NC 1.64 \R {L} \log N_H +0.951 \R {L} \R {\log N_H}^3 \NR
}
We are now in a position to bound \m {\VNm {d} _2}, just by using \Rf {d-2-dA-C1} again, and plugging in \Rf {d-2-16-H3} and \Rf {d-C1-16-H2} above.

In addition, we shall discuss the four events of failure, namely
\DispNum {q-q-2q-ar} {
\NC q
\eqsim \NC 2 \RB {q _{\Rm {iso}, B} + q _{\Rm {iso}, R}} +q _{\Rm {no}} +q _{\Rm {sp}} \NR
}
subscript respectively for restricted isometry, noise, sparsity, respectively.
Quoting previous results \Rf {1-p-1N-H3}, \Rf {P-d-eN-d4}, \Rf {1-p-ex-d's}, and \Rf {1-p-NH-H2}, we make some approximation and set
\DispNum {q-B-2e-Lp'} {
\NC q _{\Rm {iso}, B}
\lesssim \NC 2 \Ss {e} ^{-N_B /12} \NR
%
\NC q _{\Rm {iso}, R}
\lesssim \NC 2 N_H ^{-1/3} \RB {\log N_H} ^{-1/6} \NR
%
\NC q _{\Rm {no}}
\lesssim \NC N_H ^{-2} \NR
%
\NC q _{\Rm {sp}}
\lesssim \NC 2 \Ss {e} ^{-9L/\pi} \NR
}
%
If we provide the design values
\DispNum {N-B-4l-H2} {
\NC N_B
\gtrsim \NC 4 \log N_H \NR
%
\NC N_R
\gtrsim \NC 16 L \RB {\log N_H}^2 \NR
}
%
It can be seen that
\DispNum {q-p-qn-oR} {
\NC q _{\Rm {sp}}
\lesssim \NC q _{\Rm {no}} \NR
%
\NC \lesssim \NC q _{\Rm {iso}, B} \NR
%
\NC \eqsim \NC q _{\Rm {iso}, R} \NR
}
%
then we get the estimation
\DispNum {q-q-ON-13} {
\NC q
\eqsim \NC q _{\Rm {iso}, R} \NR
%
\NC \eqsim \NC q _{\Rm {iso}, B} \NR
%
\NC \eqsim \NC 2 N_H ^{-1/3} \NR
}
%

%gC1 {g-C-4p'-NH}
%dC1 {d-1-dA-C1}
%PPd-infty {P-d-4N-Nh}
%d2 {d-2-dA-C1}
%dAB2 {d-2-11-C1}

\stopsubsection
\stopsection