### 2019-11-05 presentation-B/4_simulation.tex

\Frame {Real Valued Representation}
{
\I To solve the convex optimization, we use Python package CVXPY.

\I Since it doesn't allow complex variables, we represent complex number by real part and imarinary parts.

\I For vectors, substitute
\Disp {
\NC a +b \Ss{i}
\leftrightarrow \NC \startTheMatrix
\NC a \NR
\NC b \NR
\stopTheMatrix \NR
}

\I For matrices, substitute
\Disp {
\NC a +b \Ss{i}
\leftrightarrow \NC \startTheMatrix
\NC a, \NC -b \NR
\NC b, \NC a \NR
\stopTheMatrix \NR
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Algorithm Cast as SOCP (1/3)}
{
\I Given are \m{\M {P} \in \MB {C} ^{N_Y^2 \D N_H^2}}, \m {\V {y} \in \MB {C} ^{N_Y^2}}, \m{\g > 0}.
Let \m {\T {\V {y}}} be the real representation of \m {\V {y}}, and \m {\T {\M {P}}} that of \m {\M {P}}.

\I Define the indicator vectors and matrices as
\Disp {
\NC {\V {u}}_{n_h} \SB {n_h'}
= \NC \startcases
1, \Q \MC n_h' =n_h \NR
0, \Q \NC \Rm {otherwise} \NR
\stopcases \NR
%
\NC {\T {\M {U}}}_{n_h} \SB {n_h', n_h''}
= \NC \startcases
1, \Q \MC n_h' =2n_h,\; n_h'' =2n_h \NR
1, \Q \MC n_h' =2n_h +1,\; n_h'' =2n_h +1 \NR
0, \Q \NC \Rm {otherwise} \NR
\stopcases \NR
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Algorithm Cast as SOCP (2/3)}
{
\I To fit the form of SOCP, we just have to define \m {\V {t}, \M {A}_i, \V {b}_i, \V {c}_i, d_i}, for \m {i = 0, 1, \ldots, N_H^2 -1},
\Disp{
\NC \V {t}
= \NC \startTheMatrix
\NC \V {0} \NR
\NC \V {1} \NR
\stopTheMatrix,
\Q \M {A}_i
= \startTheMatrix
\NC \T {\M {U}}_{i}, \NC \M {0} \NR
\stopTheMatrix \NR
%
\NC \V {b}_i
=\NC \V {0},
\Q \V {c}_i
= \startTheMatrix
\NC \V {0} \NR
\NC \V {u}_{i} \NR
\stopTheMatrix,
\Q d_i
= 0 \NR
}

\I Similarly, for \m {i = N_H^2, N_H^2 +1, \ldots, 2N_H^2 -1},
\Disp{
\NC \M {A}_i
= \NC \startTheMatrix
\NC -\T {\M {U}}_{i -N_H^2} \T {\M {P}}^\Adj \T {\M {P}}, \NC \M {0} \NR
\stopTheMatrix, \NR
%
\NC \V {b}_i
=\NC \T {\M {U}}_{i -N_H^2} \T {\M {P}}^\Adj \T {\V {y}},
\Q \V {c}_i
= \V {0},
\Q d_i
= \g \NR
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Algorithm Cast as SOCP (3/3)}
{
\I Now, run the Second Order Cone Programming:
\Disp{
\NC \Hat {\V {x}}
\LA \NC \startcases
\NC \Min {\V {x}'}
\MC \IP {\V {t}, \V {x}'} \NR
\NC \Rm {subject} \; \Rm {to}
\Q  \MC \VNm {\M {A} _{i} \V {x}' +\V {b} _{i}} _2
\leq \IP {\V {c}_{i}, \V {x}'} +\V {d}_i \NR
\NC \MC i 
=0, 1, \ldots, 2N_H^2 -1 \NR
\stopcases \NR
}
\I Extract channel response as real vector:
\Disp {
\NC \Hat{\V {g}}
\LA \NC \MC{R} ^{-1} \SB {\Hat{\V {x}} \DB {0 : 2N_H^2-1}} \NR
}
\I Convert back from Angular domain and output:
\Disp {
\NC \Hat {\M {H}}
\LA \NC \M {K} \Rm {vec}^{-1} \SB {\Hat {g}} \M {K}^\Adj \NR
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Performance Metric}
{
\I Bound of log of error norm:
\Disp {
\NC \h_{\Ss{th}}
\eqsim \NC \F {1} {2} \log L +\F {3} {2} \log \log N_H +2 \NR
}
\I Log absolute / relative error:
\Disp {
\NC \h_{\Ss{abs}}
=\NC \RB {\log_2 {\VNm {\V {h} -\Hat {\V {h}}}_2}} _{\Ss{avg}} \NR
%
\NC \h_{\Ss{rel}}
=\NC \F {\h_{\Ss{abs}}}
{\RB {\log_2 {\VNm {\V {h}}_2}} _{\Ss{avg}}} \NR
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Simulation Parameters}
{
\I Antennae number of channel \m {N_H = 16, 32, 48}
\I Antennae number of analog stage \m {N_R = \RB {\log N_H}^2}
\I Antennae number of digital stage \m {N_Y = \log N_H}
\I Analog beamformer quantization grid: \m {16}
\I Number of paths \m {L =4}
\I How many noise standard deviation tried: \m {12}
\I Number of repetition of each data point: \m {16}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Absolute Error, \m {N_H=32}}
{
\blank [big]
\externalfigure [medium-assorted-abs.png] [wfactor=fit, hfactor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Relative Error, \m {N_H=32}}
{
\blank [big]
\externalfigure [medium-assorted-rel.png] [wfactor=fit, hfactor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Time Taken, \m {N_H=32}}
{
\blank [big]
\externalfigure [medium-assorted-time.png] [wfactor=fit, hfactor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Absolute Error, \m {N_H=48}}
{
\blank [big]
\externalfigure [medium-assorted-abs.png] [wfactor=fit, hfactor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Relative Error, \m {N_H=48}}
{
\blank [big]
\externalfigure [medium-assorted-rel.png] [wfactor=fit, hfactor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Time Taken, \m {N_H=48}}
{
\blank [big]
\externalfigure [medium-assorted-time.png] [wfactor=fit, hfactor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Discussion}
{

\I \m {\eta _{\Ss {th}}} is in correct order of magnitude but not tight when \m {\s \gg 1}

\I The smaller the noise std., the smaller the error

\I It is \It {not} true that the smaller \m {\g _{\Ss {DS}}}, the smaller the error

\I We have reasonable results, but exceedingly high complexity

\I Have to do more experiments to smooth out the lines

\I The requirement \m {N_R \gtrsim 16 L \RB {\log N_H}^2} is demanding
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %