### 2020-02-27 paper/notation.tex

\starttitle [title={Notation}]

\starttable[|l|l|l|] \HL
\NC Expression \VL Meaning or Definition \VL More Explanation \SR \HL
\NC \m {\RB {1+3} \RB {2+4}} \VL round parentheses \VL used for precedence \AR \HL
\NC \m {f \SB {x}} \VL square brackets \VL used for arguments or parameter \AR \HL
\NC \m {\CB {2, 3, 5, 7}} \VL curly braces \VL used for sets \AR \HL
\NC \m {\Nm {a}} \VL magnitude \VL for real or complex numbers \AR \HL
\NC \m {\MF {Re} \SB {a}} \VL real part \VL for complex numbers \AR \HL
\NC \m {\MF {Im} \SB {a}} \VL imaginary part \VL for complex numbers \AR \HL
\NC \m {\V {a}} \VL vector \VL \AR \HL
\NC \m {\M {A}} \VL matrix \VL \AR \HL
\NC \m {\V {1}} \VL all-1 vector \VL dimension will be specified \AR \HL
\NC \m {\V {0}} \VL all-0 vector \VL dimension will be specified \AR \HL
\NC \m {\V {a} _{\SB {n}}} \VL \m {n}-th component of \m {\V {a}} \VL \AR \HL
\NC \m {\M {A} _{\SB {m,n}}} \VL \m {m,n}-th entry of \m {\M {A}} \VL \AR \HL
\NC \m {\V {a} _{\SB {0}}} \VL the first component of \m {\V {a}} \VL  \AR \HL
\NC \m {\M {A} _{\SB {0,0}}} \VL the upper-left entry of \m {\V {A}} \VL  \AR \HL
\NC \m {\MB {K}} \VL either \m {\MB {R}} or \m {\MB {C}} \VL  \AR \HL
\NC \m {\MB {K} ^{N}} \VL Hilbert space \m {\MB {K}^N} over \m {\MB {K}} \VL  \AR \HL
\NC \m {\MB {K} ^{M \D N}} \VL collection of \m {M} by \m {N} matrices \VL  \AR \HL
\NC \m {m\; \Rm {Mod}\; n} \VL remainder of \m {m} divided by \m {n} \VL with \m {m, n \geq 0} and \m {0 \leq m\; \Rm {Mod}\; n <n} \AR \HL
\NC \m {\VNm {\V {a}} _p} \VL \m {\ell_p}-norm of \m {\V {a}} \VL \m {p >0} \AR \HL
\NC \m {\VNm {\V {a}} _\infty} \VL the maximum-magnitude \VL  \AR
\NC  \VL component of \m {\V {a}} \VL  \AR \HL
\NC \m {\MB {P} \SB {A}} \VL probability of event \m {A} \VL  \AR \HL
\NC \m {\MB {E} \SB {X}} \VL expectation of random variable \m {X} \VL  \AR \HL
\NC \m {\Ss {Var} \SB {X}} \VL variance of random variable \m {X} \VL  \AR \HL
\stoptable

\stopchapter