### 2020-03-24 presentation/1_introduction.tex
Update slides

\blank [big, force]

\Title {\TitleText}
\Subtitle {Numerical Experiments}
\blank [big]

\Subtitle {\AuthorText}
\blank [big]

\Subsubtitle {\DateText}

\page [yes]
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Organization}
{
\I Problem Setting

\I Proposed Method

\I Bound for the Error Norm

\I Numerical Experiments

\I Conclusion and References
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Background}
{
\I Multiple-input multiple-output (MIMO) communication system has been proposed as part of the 5G specification, but its hardware overhead increases complexity and power consumption.

\I The design of new algorithms addressing these issues requires channel information to be known.

\I Estimation of MIMO channels also results in high complexity, due to the large number of antennae.

\I Meanwhile, millimeter wave (mm-Wave) channels often exhibit sparse properties, which we exploit to save for some measurements.
}