### 2020-03-24 presentation/2_setting.tex

\Frame {Channel Model}
{
\I Consider the uniform linear array
\Disp {
\NC \V {a} \SB {\psi'}
= \NC \F {1}{\R {N_H}} \sum_{n=0}^{N-1} \Ss {e}^{n \psi' \Ss {i}} \V {u}_n
\in \MB {C} ^{N_H} \NR
}

\I Consider the channel representation
\Disp {
\NC \m {H}
=\NC \sum_{l=0} ^{L-1}
\a_l
\V {a} \SB {\f_l'}
\V {a} \SB {\th_l'}^\Adj
\in \MB {C} ^{N_H \D N_H} \NR
}

\I where \m {L} is the number of paths, \m {\f_l'} is related to the angles of arrival, \m {\th_l'} is related to the angles of arrival.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Precoders and Combiners}
{
\I Tx: Digital precoder \m {\M {F}_B \in \NC \MB {C} ^{N_R \D N_Y}}

\I Tx: Analog precoder \m {\NC \M {F}_R \in \NC \MB {C} ^{N_H \D N_R}}

\I Rx: Digital combiner \m {\M {W}_B \in \NC \MB {C} ^{N_R \D N_H}}

\I Rx: Analog combiner \m {\M {W}_R \in \NC \MB {C} ^{N_Y \D N_R}}

\I With the unity magnitude constraint
\Disp {
\NC \Nm {\M {F}_R \SB {n_H, n_R}}
= \NC 1,
\Q \Nm {\M {W}_R \SB {n_R, n_H}}
= 1, \NR
}

\I Under the assumption
\Disp {
\NC N_H \gg \NC N_R \gg N_Y \NR
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Effective Channel}
{
\I Consider Complex Standard Normal \m {\m {Z} \in \NC \MB {C} ^{N_H \D N_H}}

\I Introduce the effective channel
\Disp {
\NC \m {Y}
:=\m {W}_B \m {W}_R \RB {\m {H} \m {F}_R \m {F}_B +\m {Z}} \NR
}

\I Their dimension is summarized for convenience:
\Disp {
\NC \MB {C}^{N_Y}
\underset {\m {F}_B} {\mapsto} \NC \MB {C}^{N_R}
\underset {\m {F}_R} {\mapsto} \MB {C}^{N_H}
\underset {\m {H}} {\mapsto} \MB {C}^{N_H}
\underset {\m {W}_R} {\mapsto} \MB {C}^{N_R}
\underset {\m {W}_B} {\mapsto} \MB {C}^{N_Y} \NR
}

\I Problem (a): How do we recover \m {\M {H}} with knowledge of \m {\M {Y}}?

\I Problem (b): How do we design \m {\M {W}_R}, \m {\M {W}_B}, \m {\M {F}_R}, and \m {\M {F}_B}?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Transform to the Angular Domain}
{
\I Introduce the DFT matrix \m {\M {K} \in \NC \MB {C} ^{N_H \D N_H}}, and write
\Disp {
\NC \V {y}
:= \NC \Rm {vec} \SB {\M {Y}},
\Q \M {P}
:= \RB {\M {F}_B^\Tr \M {F}_R^\Tr \m {K}^\ast} \otimes \RB {\M {W}_B \M {W}_R \M {K}}, \NR
\NC \V {g}
:= \NC \Rm {vec} \SB {\M {K}^\Adj \M {H} \M {K}},
\Q \V {z}
:= \Rm {vec} \SB {\M {Z}} \NR
}

\I Then we have a linear equation
\Disp {
\NC \V {y}
=\NC \M {P} \V {g} +\V {z} \NR
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Compressive Sensing}
{
\I In our case, we have to essentially invert an underdetermined system, and usually, that is not well defined.

\I Consider a simpler case: \m {\V {y} = \M {A} \V {x} + \V {z}}, where \m {\M {A} \in \MB {R} ^{N_m \D N_p}}, \m {\V {y} \in \MB {C} ^{N_m}}.

\I If \m {\V {y} = \M {A} \V {x}_1 = \M {A} \V {x}_2}
\Disp {
\NC \V {x}_1 - \V {x}_2 \in \Rm {null}\; \M {A} \NR
}
Then \m {\V {x}_1, \V {x}_2} cannot be recovered.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Restricted Isometry Property}
{
\I However, Candès and Tao shows that when the signal is sparse, few measurements may be sufficient for the reconstruction, with vanishing error norm, in overwhelming probability.

\I \m {\V {g}} is called \Emph {\m {s}-sparse}, if only \m {s} of its components are nonzero.

\I We say that \m {\m {P}} satisfies the \Emph {restricted isometry property} (RIP) of sparsity \m {s} with respect to \m {0 < \d_s < 1}, if, for all \m {s}-sparse \m {\V {g}},
\Disp {
\NC \RB {1-\d_s} \VNm {\V {g}} _2^2
\leq \NC \VNm {\m {P} \V {g}} _2^2
\leq \RB {1+\d_s} \VNm {\V {g}} _2^2
}

\I Verbally, \m {\m {P}} is \Emph {almost unitary} up to relative error \m {\d_s}.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Dantzig Selector for Sparse Vectors}
{
\I Let \m {\V {x}} be \m {S}-sparse.
If \m {\g_{\Ss {DS}} = \R {2 \log N_p}}, then
\Disp {
\NC \Hat {\V {x}}
\LA \NC \startcases
\NC \Min {\V {x}'}  \MC \VNm {\V {x}'} _1 \NR
\NC \Rm {subject} \; \Rm {to} \Q \MC \VNm {\M {A}^\intercal \RB {\V {y} -\M {A} \V {x}'}} _\infty \leq \g_{\Ss {DS}} \NR
\stopcases \NR
}
with
\Disp {
\NC \VNm {\V {x} - \hat {\V {x}}} _2 ^2
\leq \NC 32 S \sigma^2 \log N_p \NR
}
For high probability.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Proposed Method}
{
\I Back to our problem: Input \m {\M {P} \in \MB {C} ^{N_Y^2 \D N_H^2}}, \m {\V {y} \in \MB {C} ^{N_H^2}}, \m {\g_{\Ss {DS}} \geq 0}.

\I Calculate
\Disp {
\NC \Hat {\V {g}}
\LA \NC \startcases
\NC \Min {\V {g}'}  \MC \VNm {\V {g}'} _1 \NR
\NC \Rm {subject} \; \Rm {to} \Q \MC \VNm {\M {P}^\Adj \RB {\V {y} -\M {P} \V {g}'}} _\infty \leq \g_{\Ss {DS}} \NR
\stopcases \NR
}

\I Output \m {\Hat {\M {H}} \LA \M {K} \Rm {vec}^{-1} \SB {\Hat {g}} \M {K}^\Adj}
}