### 2020-03-24 presentation/4_simulation.tex

% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Linear Programming}
{
\I Set \m {\V {1} = \IP {1, \dots, 1}}.
\I Calculate
\Disp {
\NC \Hat {\V {g}}, \Hat {\V {f}}
\LA \NC \startcases
   \NC \Min {\V {g}', \V {f}'}  \MC \IP {\V {1}, \V {f}'} \NR
   \NC \Rm {subject} \; \Rm {to} \MC \V {g}' \leq \V {f}' \NR
   \NC \MC - \V {g}' \leq \V {f}' \NR
   \NC \MC \M {P}^\Adj \M {P} \V {g}' \leq \M {P}^\Adj \V {y} + \g_{\Ss {DS}} \V {1} \NR
   \NC \MC - \M {P}^\Adj \M {P} \V {g}' \leq - \M {P}^\Adj \V {y} + \g_{\Ss {DS}} \V {1} \NR
\stopcases \NR
}
\I Output \m {\Hat {\M {H}} \LA \M {K} \Rm {vec}^{-1} \SB {\Hat {g}} \M {K}^\Adj}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Two-Stage Estimation (1/2)}
{
\I Denote
\Disp {
\NC N_0 = \NC 2 N_H^2 \NR
\NC N_1 = \NC 2 N_H \NR
\NC N_2 = \NC 2 \lfloor \Rm {min} \SB {L \log \SB {N_H^2}, N_H} \rfloor \NR
}

\I Apply DS to \m {\V {y}, \M {P}} to get \m {\Hat {\V {g}}_0}.

\I Choose the \m {N_1} largest component of \m {\Hat {\V {g}}_0} and call it \m {\V {g}_1}.
Choose corresponding columns of \m {\M {P}} and call it \m {\M {P} _1}.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Two-Stage Estimation (2/2)}
{
\I Apply DS to \m {\V {y}, \M {P} _1} to get \m {\Hat {\V {g}}_1}.

\I Choose the \m {N_2} largest component of \m {\Hat {\V {g}}_1} and call it \m {\V {g}_2}.
Choose corresponding columns of \m {\M {P}_1} and call it \m {\M {P} _2}.

\I Apply pseudoinverse (linear square) to \m {\V {y}, \M {P} _2} to get \m {\Hat {\V {g}}_2}.

\I Assign corresponding components of \m {\Hat {\V {g}}_2} in \m {\Hat {\V {g}}}, and pad \m {0} otherwise.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Performance Metric}
{
\I With some approximations, I suggest the following metric for spectral efficiency:
\I Denote \m {E := \Hat {H} - H}
\I Observe
\Disp {
\NC \VNm {\RB {\M {H} + \M {E}} \RB {\M {H} + \M {E}} ^\Adj - \M {H} \M {H} ^\Adj} \V {x}
\lesssim 2 \VNm {\M {E}} _2 \VNm {\M {H}} _2 \VNm {\V {x}} \NR
}
\I Define the spectral efficiency to be
\Disp {
\NC \log \det \SB {I + \RB {\s + \F {2 \VNm {\M {E}} _2 \VNm {\M {H}} _2} {N_H ^{3/2}}} ^{-1} \M {H} \M {H} ^ \Adj} \NR
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Theoretical Metric}
{
\I For the theoretical error norm,
\Disp {
\NC \log d_{\Rm {th}}
\lesssim \NC \F {1} {2} \log L +\F {3} {2} \log \log N_H \NR
}
\I Define the theoretical spectral efficiency to be
\Disp {
\NC \log \det \SB {I + \RB {\s + \F {2 d_{\Rm {th}} \VNm {\M {H}} _2} {N_H ^{3/2}}} ^{-1} \M {H} \M {H} ^ \Adj} \NR
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Simulation Parameters}
{
\I Antennae number of channel \m {N_H = 16, 20, 24}
\I Antennae number of analog stage \m {N_R = \lfloor \R {N_H N_Y} \rfloor}
\I Antennae number of digital stage \m {N_Y = \lfloor N_H/3 \rfloor}
\I Number of paths \m {L =3}
\I Analog beamformer quantization grid: \m {16}
\I How many noise standard deviation tried: \m {7}
\I Number of repetition of each data point: \m {12}
\I Solved by Python CVXPY
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Spectral Efficiency, \m {N_H=16}}
{
\blank [big]
\externalfigure [assorted-small-efficiency.png] [factor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Time Taken, \m {N_H=16}}
{
\blank [big]
\externalfigure [assorted-small-time.png] [factor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Spectral Efficiency, \m {N_H=20}}
{
\blank [big]
\externalfigure [assorted-medium-efficiency.png] [factor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Time Taken, \m {N_H=20}}
{
\blank [big]
\externalfigure [assorted-medium-time.png] [factor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Spectral Efficiency, \m {N_H=24}}
{
\blank [big]
\externalfigure [assorted-big-efficiency.png] [factor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Time Taken, \m {N_H=24}}
{
\blank [big]
\externalfigure [assorted-big-time.png] [factor=fit]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Discussion}
{
\I DS is slightly better than others when \m {\s \gg 1}

\I \m {\eta _{\Ss {th}}} is in correct order of magnitude but not tight when \m {\s \gg 1}

\I The smaller the noise, the smaller the error

\I It is \It {not} true that the smaller \m {\g _{\Ss {DS}}}, the smaller the error

\I We have exceedingly high complexity than OMP

\I The requirement that \m {N_H \gg N_Y} is demanding
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Complexity (1/2)}
{
\I In general, the complexity of convex optimization is hard to obtain.

\I However, there exist expressions for number of Newton steps in the course of optimization.
It is \m {\MC {O} \SB {\VNm {\V {g} _0 - \V {g} ^\star} _1}}, where \m {\VNm {\V {g} _0} _1} is the starting value, \m {\V {g} ^\star} the true value.
Assume that is \m {\MC {O} \SB {\VNm {\M {P} ^\Adj \V {z}} _1}}, which is at most \m {\MC {O} \SB {N_H^2}}.

\I Moreover, every Newton step requires a matrix inversion of dimension \m {N_H^2}, with at least \m {\MC {O} \SB {N_H^4}}.
This gives \m {\MC {O} \SB {N_H^6}}.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Complexity (2/2)}
{
\I On the other hand, it is pointed out that Lasso (least absolute shrinkage and selection operator) has similar complexity with Least Angle regression, which has complexity \m {\MC {O} \SB {N_H^6}}.
Since DS and Lasso have similar form, we expect DS has similar complexity.

\I Clearly, the complexity of OMP is at most \m {\MC {O} \SB {N_H^2}}, so much smaller than DS.
}