### 2020-03-24 presentation/5_conclusion.tex

\Frame {Conclusion}
{
\I We propose a method to estimate MIMO mm-Wave channel \m {\M {H}} by exploiting its sparsity, and employing Dantzig Selector (DS).

\I We modified DS for present setting, for the almost-sparse condition in the angular domain, using the effective beamformer \m {\M {P}}, and concerning complex vectors.

\I We have a bound of the error norm with respect to parameters \m {N_H}, \m {N_Y}, and \m{L}, and investigated the asymptotic behavior of the successful probability.

\I We adopt linear programming to reduce complexity.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Future Work}
{
\I Is it possible to resort to less precision to reduce complexity, but retaining the performance?

\I Run for bigger values of \m {N_H}, \m {N_Y} and repeat for enough times to observe a definite result.

\I I want to know why DS is not as good as expected, and to provide discussion, before closing the case.

\I The proof section need some rewriting, but it is not big deal.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {References}
{
{\tfx
\I S Boyd and L Vandenberghe, \It {Convex optimization}, Cambridge U Press, 2004.

\I E Candès and J Romberg, \Quote {\m {\ell_1}-magic: Recovery of Sparse Signals via Convex Programming}, retrieved online, 2005.

\I E Candès and T Tao, \Quote {The Dantzig Selector: Statistical estimation when \m {p} is much larger than \m {n}}, \It {The Annals of Statistics} 35(6), 2007.

\I B Efron and T Hastie, \Quote {Least angle regression}, \Quote {The Annals of statistics}, 32(2), 2004.

\I J Lee, G-T Lee, and H Yong, \Quote {Channel estimation via orthogonal matching pursuit for hybrid MIMO systems in millimeter wave communications}, \It {IEEE Transactions on Communications}, 64(6), 2016.
}
}