### 2020-06-09 presentation/5_conclusion.tex

\Frame {Conclusion}
{
\I We propose a method to estimate MIMO mm-Wave channel \m {\M {H}} by modifying Dantzig Selector (DS), exploiting the sparsity inherent in the problem.

\I We argue for its validity, for the almost-sparse condition in the angular domain, using the effective beamformer \m {\M {P}}, and concerning complex vectors.

\I We propose a bound of the error norm with respect to parameters \m {N_H}, \m {N_Y}, and \m{L}, and investigated the asymptotic behavior of the successful probability.

\I We convert the problem into a linear program to reduce complexity, and run a two-stage version to estimate the position of components.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Future Work}
{
\I Is it possible to reduce complexity, while retaining the performance?

\I Is the trend the same for bigger values of \m {N_H}?

\I Why DS is not as good as expected, and is there any mistakes in the proof of error bound?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {References (1/2)}
{
{\tfx
\I W U Bajwa, J Haupt, A M Sayeed, and R Nowak, \Quote {Compressed channel sensing: A new approach to estimating sparse multipath channels}, \It {Proceedings of the IEEE}, 98(6), 2010

\I S Boyd and L Vandenberghe, \It {Convex optimization}, Cambridge U Press, 2004.

\I E Candès and J Romberg, \Quote {\m {\ell_1}-magic: Recovery of Sparse Signals via Convex Programming}, retrieved online, 2005.

\I E Candès and T Tao, \Quote {The Dantzig Selector: Statistical estimation when \m {p} is much larger than \m {n}}, \It {The Annals of Statistics} 35(6), 2007.
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {References (2/2)}
{
{\tfx
\I E J Candès and T Tao, \Quote {Decoding by Linear Programming}, \It {IEEE Transactions on Information Theory}, 51(12), 2005

\I M P Friedlander and M A Saunders, \Quote {Discussion: the Dantzig selector: statistical estimation when p is much larger than n}, \It {The Annals of Statistics}, 35(6), 2007

\I J Lee, G-T Lee, and H Yong, \Quote {Channel estimation via orthogonal matching pursuit for hybrid MIMO systems in millimeter wave communications}, \It {IEEE Transactions on Communications}, 64(6), 2016.

\I J A Tropp and A C Gilbert, \Quote {Signal recovery from random measurements via orthogonal matching pursuit}, \It {IEEE Transactions on information theory}, 53(12), 2007.
}
}