### 2020-08-17 treatise/3_3_generalizing_the_dantzig_selector_bound.tex

\startsection [title={Generalizing the Dantzig Selector Bound}]
\startsubsection [title={Technical Propositions in Complex Case}]

The third and last part of our task is to plug in the almost-sparsity found in Section 3.1, and plug in the probablity found in Section 3.2.
Set for short
%
\DispNum {d:d:gg:gg} {
\NC \V {d} 
= \NC \Hat {\V {g}} - \V {g} \NR
}
%
And our task is to bound the error norm \m {\VNm {\V {d}} _2}.

We argue that \m {\V {g}} can essentially be seen as sparse, and for that purpose we approximate irrelevent terms.
For example, we generously bound \m {\d_{S}} by \m {1}.
And, as shall be seen, it suffices to set
%
\DispNum {g':g':2N:Nh} {
\NC \g
= \NC \R {4 \log N_H} \NR
}
%
The propositions below are taken from \cite [CaT07] and modified slightly in accordance to our settings, except possibly the generalization to complex vector.
%
\Result
{Proposition}
{
%
\DispNum {d:1:dA:C1} {
\NC \VNm {\V {d} _{\SB{\MS {C}}}} _1
\leq \NC \VNm {\V {d} _{\SB{\MS {A}}}} _1
+2\VNm {\V {g} _{\SB{\MS {C}}}} _1 \NR
}
}
%
To show this, observe that with triangle inequality applied respectively on \m {\MS {A}} and \m {\MS {C}}, while recalling that \m {\V {g} =\V {g} _{\SB{\MS {A}}}},
%
\DispNum {g:1:gd:g1} {
\NC \VNm {\V {g}} _1
- \VNm {\V {d} _{\SB{\MS {A}}}} _1
+ \VNm {\V {d} _{\SB{\MS {C}}}} _1
- \VNm {\V {g} _{\SB{\MS {C}}}} _1
\leq \NC \VNm {\V {g} + \V {d}} _1 \NR
%
\NC =\NC \VNm {\hat {\V {g}}} _1 \NR
}
%
Also, by construction that \m {\hat {g}} minimizes the \m {\ell_1}-norm,
%
\DispNum {g:1:g1:g1} {
\NC \VNm {\hat {\V {g}}} _1
\leq \NC \VNm {\V {g}} _1 \NR
}
as desired.

We turn our attention to the next result, and fix an instance of \m {\M {P}} and instance of \m {\V {z}}.
We claim that
%
\DispNum {E:h:2N:Nh} {
\NC \MB {E} \SB {\Nm {\IP {\V {z} , \M {P} _{\SB {:, n_h}}}}}
\leq \NC  2 \R {2 \log N_H} \NR
%
\NC n_h 
=\NC 1, \ldots, N_H^2. \NR
}
%
Start with the fact that \m {\M {P} _{\SB {:, n_g}}} has unity \m {\ell_2}-norm, and on the randomness of \m {\V {z}}.
Now, the quantity \m {\Nm {\IP {\V {z} , \M {P} _{\SB {:, n_h}}}}} observes Standard Normal.
To bound the failure probability, recall the tail bound for \m {Q} function
%
\DispNum {Q:x:12:22} {
\NC Q\SB {x'}
\leq \NC \F {1} {2} \Ss {e}^{-x'^2 /2} \NR
}
%
Particularly, for \m {x' =\R {4 \log N_H}}, this gives the failure probability
%
\DispNum {Q:h:1N:Nh} {
\NC Q\SB {\R {4 \log N_H}}
=\NC \F {1} {N_H^2}. \NR
}
%
\Result
{Proposition}
{
If \m {\M {P}} is \m {\d_S}-RIP,
%
\DispNum {P:d:4N:Nh} {
\NC \VNm {\M {P}^\Adj \M {P} \V {d}} _\infty
\leq \NC  4 \R {2 \log N_H} \NR
}
%
holds for probability \m {p} 
%
\DispNum {1:p:NH:H2} {
\NC 1 -p
\leq \NC N_H^{-2}. \NR
}
}
%
To show this, by definition
%
\DispNum {z:y:Pg:ny} {
\NC \IP {\V {z} - \RB {\V {y} - \M {P} \hat {\V {g}}}, \M {P} _{\SB {:,n_y}}}
= \NC \IP {\M {P} \hat {\V {g}} - \M {P} \V {g}, \M {P} _{\SB {:,n_y}}} \NR
%
\NC = \NC \IP {\M {P} \V {d}, \M {P} _{\SB {:,n_y}}} \NR
}
%
By construction
%
\DispNum {P:g:Py:Nh} {
\NC \VNm {\M {P} _{\SB {:,n_y}}^\Adj \RB {\V {y} - \M {P} \hat {\V {g}}}} _\infty
\leq \NC \VNm {\M {P}^\Adj \RB {\V {y} - \M {P} \hat {\V {g}}}} _\infty \NR
%
\NC \leq \NC 2\R {2 \log N_H} \NR
}
%
By triangle inequality, together with \Rf {Q:h:1N:Nh}, we obtain the desired result.
%
\DispNum {P:d:zP:Nh} {
\NC \VNm {\M {P} _{\SB {:,n_y}}^\Adj \M {P} \V {d}} _\infty
\leq \NC \VNm {\IP {\V {z}, \M {P} _{\SB {:,n_y}}}} _\infty
+ \IP {\V {y} - \M {P} \hat {\V {g}}, \M {P} _{\SB {:,n_y}}} \NR
%
\NC \leq \NC 2 \R {2 \log N_H} +2 \R {2 \log N_H} \NR
}
%
Lastly, the following two propositions are taken from \cite [CaT07], Lemma 1.
The generalization from real vector spaces to complex spaces is completely straightforward, while meaning of magnitude and inner product has to be changed.
One of the quoted propositions reads
%
\DispNum {d:2:dA:12} {
\NC \VNm {\V {d}} _2^2
\leq \NC \VNm {\V {d} _{\SB {\MS {AB}}}} _2^2 + \F {1} {S} \VNm {\V {d} _{\SB{\MS {C}}}} _1^2 \NR
}
%
By power mean inequality, we get a statement cruder than \Rf {d:2:dA:12}, as below.
%
\Result
{Proposition}
{
%
\DispNum {d:2:dA:C1} {
\NC \VNm {\V {d}} _2
\leq \NC \VNm {\V {d} _{\SB {\MS {AB}}}} _2 + \F {1} {\R {S}} \VNm {\V {d} _{\SB{\MS {C}}}} _1 \NR
}
}
%
The other quoted proposition reads
%
\DispNum {d:2:11:C1} {
\NC \VNm {\V {d} _{\SB{\MS {AB}}}} _2
\leq \NC \F {1} {1- \d_{2S}} \VNm {P _{\SB {\MS {AB}}}^\Tr P d} _2
+ \F {\T {\d}_{S, 2S}} {\RB {1- \d_{2S}} \R {S}} \VNm {d_{\SB{\MS {C}}}} _1 \NR
}
%
Here \m {\T {\d} _{s_1, s_2}} is called the \m {s_1,s_2}-restricted isometry constant, and is defined in \cite [Can05] (which uses \m {\th}), as follows.
Let \m {\M {Q}'} be fixed.
For all \m {s_1}-sparse \m {\V {x}_1}, for all \m {s_2}-sparse \m {\V {x}_2}, whose nonzero-positions are disjoint,
%
\DispNum {d:2:in:x2} {
\NC \T {\d} _{s_1, s_2}
= \NC \inf \CB {
\d':\; \Nm {\IP {\M {Q}' \V {x}_1, \M {Q}' \V {x}_2}}
   \leq \d' \VNm {\V {x}_1} \VNm {\V {x}_2}
} \NR
}
%
Here we use a result in \cite [Can05] that relates \m {\T {\d} _{s_1, s_2}} to \m {\d_s}, that is
\DispNum {d:2:ds:s2} {
\NC \T {\d} _{s_1, s_2}
\leq \NC \d_{s_1+s_2} \NR
}
And we need another crude bound to relate \m {\d_{2S}} to \m {\d_{S}}.
By triangle inequality, by power-mean inequality,
\DispNum {Q:x:Qx:x2} {
\NC \VNm {Q' \RB {\V {x}_1 + \V {x}_2}}
%
\leq \NC \VNm {Q' \V {x}_1}
+ \VNm {Q' \V {x}_2} \NR
%
\NC \leq \NC \RB {1 + \d _{s'}} \RB {\VNm {\V {x}_1} + \VNm {\V {x}_2}} \NR
%
\NC \leq \NC \RB {1 + \d _{s'}} \D \R{2} \VNm {\V {x}_1 + \V {x}_2} \NR
}
Therefore
\DispNum {d:s:2s:2s} {
\NC \d_{2s'}
\leq \NC \R {2} \d_{s'} \NR
}
By the same token,
%
\Result
{Proposition}
{
If \m {\M {Q}'} has \m {\d_{s'}}-RIP, then, for \m {M=1,2,3,\dots} so that \m {\d _{Ms'}} is meaningful,
%
\DispNum {d:s:Md':d's} {
\NC \d_{Ms'}
\leq \NC \R {M} \d_{s'} \NR
}
}
%
In particular,
%
\DispNum {d':S:Ll:d's} {
\NC \d_{S}
\leq \NC \R {L} \RB {\log N_H} ^{1/4} \d_{s} \NR
}
%
So \Rf {d:2:11:C1} becomes
%
\Result
{Proposition}
{
If \m {\M {P}} is \m {\d_S}-RIP,
%
\DispNum {d:2:11:C1} {
\NC \VNm {\V {d} _{\SB{\MS {AB}}}} _2
\leq \NC \F {1} {1- \R{2} \d_{S}} \VNm {P _{\SB {\MS {AB}}}^\Tr P d} _2
+ \F {\R{3} \d_{S}} {\RB {1- \R{2} \d_{S}} \R {S}} \VNm {d_{\SB{\MS {C}}}} _1 \NR
}
}
\stopsubsection

\startsubsection [title={Main Result}]

We are going to combine previous propositions and show the main result.
For simplicity, set \m {S = Ls^2}, thus
%
\DispNum {S:S:4N:NH} {
\NC S
=\NC L \log N_H \NR
}
%
And, for concreteness,
%
\DispNum {d':S:18:18} {
\NC \d_S
\leq \NC \F {1} {8} \NR
}
%
\Result
{Theorem}
{
Let \m {\V {y}}, \m {\M {P}}, \m {\V {g}}, \m {\hat {\V {g}}}, \m {\V {d}} be defined as above.
Then the bound
%
\DispNum {d:2:OL:H3} {
\NC \VNm {\V {d}} _2
\leq \NC \MS {O} \RB {\R {L} \R {\log N_H}^3} \NR
}
%
holds for probability \m {p}, with
%
\DispNum {1:p:ON:13} {
\NC 1 -p
=\NC \MS {O} \SB {N_H ^{-1/3}} \NR
}
}
To prove this, we simplify the expression for \m {\VNm {\V {d} _{\SB{\MS {AB}}}} _2} in \Rf {d:2:11:C1} and the one for \m {\VNm {\V {d} _{\SB{\MS {C}}}} _1} in \Rf {d:1:dA:C1}.
In the below, we take the freedom to substitute rough numbers.

By the definition of truncation, by \m {\ell_p} norm inequality, by \Rf {P:d:4N:Nh},
%
\DispNum {P:2:PP:NH} {
\NC \VNm {\M {P} _{\SB {\MS {AB}}}^\Tr \M {P} \V {d}} _2
\leq \NC \VNm {\M {P}^\Tr \M {P} \V {d}} _2 \NR
%
\NC \leq \NC \R {S} \VNm {\M {P}^\Tr \M {P} \V {d}} _\infty \NR
%
\NC \leq \NC 5.65 \R {\log N_H} \NR
}
%
By \m {\ell_p}-norm inequality, by the definition of truncation, by \Rf {d:2:11:C1}, and by \Rf {P:2:PP:NH} just above,
%
\DispNum {d:1:Sd:C1} {
\NC \VNm {\V {d} _{\SB{\MS {A}}}} _1
\leq \NC \R {S} \VNm {\V {d} _{\SB{\MS {A}}}} _2 \NR
%
\NC \leq \NC \R {S} \VNm {\V {d} _{\SB{\MS {AB}}}} _2 \NR
%
\NC \leq \NC 1.21 L \R {\log N_H}^3
+0.262 \VNm {\V {d} _{\SB{\MS {C}}}} _1 \NR
}
%
And \Rf {a:2:4N:NH} gives, after arrangement,
%
\DispNum {d:A1:16:H2} {
\NC \VNm {\V {d} _{\SB{\MS {A}}}} _1
%
\leq \NC 1.64 L \R {\log N_H}^3 +0.951 L \RB {\log N_H}^2 \NR
}
%
Then, substituting \Rf {g:C:4p':NH} and \Rf {d:A1:16:H2} above into \Rf {d:1:dA:C1}, we get
\DispNum {d:C1:16:H2} {
\NC \VNm {\V {d} _{\SB{\MS {C}}}} _1
%
\leq \NC 1.64 L \R {\log N_H}^3 +3.61 L \RB {\log N_H}^2 \NR
}
%
And \Rf {d:2:11:C1} becomes
%
\DispNum {d:2:16:H3} {
\NC \VNm {\V {d} _{\SB{\MS {AB}}}} _2
%
\leq \NC 1.64 \R {L} \log N_H +0.951 \R {L} \R {\log N_H}^3 \NR
}
We are now in a position to bound \m {\VNm {d} _2}, just by using \Rf {d:2:dA:C1} again, and plugging in \Rf {d:2:16:H3} and \Rf {d:C1:16:H2} above.
This gives
\DispNum {d:2:42:NH} {
\NC \VNm {\V {d}} _2
\leq \NC 3.29 \R {L} \log N_H +4.56 \R {L} \R {\log N_H}^3 \NR
}

Furthermore, we discuss the four events of failure, as follows.
Consider failure probabilities \m {q _{\Rm {iso}, B}}, \m {q _{\Rm {iso}, R}}, \m {q _{\Rm {no}}}, \m {q _{\Rm {sp}}}, respectively for restricted isometry, noise, sparsity.
If all of them are small, a naive application of union bound suggests
\DispNum {q:q:2q:ar} {
\NC q
\eqsim \NC 2 \RB {q _{\Rm {iso}, B} + q _{\Rm {iso}, R}} +q _{\Rm {no}} +q _{\Rm {sp}} \NR
}
Quoting previous results \Rf {1:p:2e:s4}, \Rf {1:p:ex:d's}, \Rf {1:p:ex:d's}, and \Rf {1:p:NH:H2}, we make some approximation and set
\DispNum {q:B:2e:Lp'} {
\NC q _{\Rm {iso}, B}
\lesssim \NC 2 \Ss {e} ^{-N_B /12} \NR
%
\NC q _{\Rm {iso}, R}
\lesssim \NC 2 N_H ^{-1/3} \RB {\log N_H} ^{-1/6} \NR
%
\NC q _{\Rm {no}}
\lesssim \NC N_H ^{-2} \NR
%
\NC q _{\Rm {sp}}
\lesssim \NC 2 \Ss {e} ^{-9L/\pi} \NR
}
%
If we provide the design values
\DispNum {N:B:4l:H2} {
\NC N_B
\gtrsim \NC 4 \log N_H \NR
%
\NC N_R
\gtrsim \NC 16 L \RB {\log N_H}^2 \NR
}
%
It can be seen that
\DispNum {q:p:qn:oR} {
\NC q _{\Rm {sp}}
\lesssim \NC q _{\Rm {no}} \NR
%
\NC \lesssim \NC q _{\Rm {iso}, B} \NR
%
\NC \eqsim \NC q _{\Rm {iso}, R} \NR
}
%
then we get the estimation
\DispNum {q:q:ON:13} {
\NC q
\eqsim \NC 2 N_H ^{-1/3} \NR
}

\stopsubsection
\stopsection