### 2020-08-17 treatise/conclusion.tex

\startchapter [title={Conclusion}]

This treatise studied the effective estimation of the millimeter wave, multiple-input, multiple-output (mm-wave MIMO) channel.
We adopts the Dantzig Selector (DS) rather than Orthogonal Matching Pursuit (OMP), which is most commonly used in literature, as the estimation algorithm, therefore exploiting the sparsity of the problem in spacial Fourier domain.
We bounded the expected square error following the spirit of Candès and Tao \cite [CaT07].
Towards that end, we characterized the spacial sparsity, confirmed the restricted isometry of measurement matrix \m {\M {P}}, and justified the generalization for complex vectors.
The bound assumes a fairly general setting, and is expressed in terms of physical parameters, such as the number of path \m {L} and channel dimension \m {N_H}.

Simulation is done, and DS's performance is compared to OMP, Lasso, and LS.
DS performs slightly better than others, but gives much higher complexity, agreeing our analysis.
Put concisely, no algorithm is a universal solution, as they trade off precision for time and space complexity, or conversely, among other things.
We hope that our inquiry of DS suggests a welcomed alternative added to the present literature of channel estimation.

\stopchapter