### 2020-10-21 presentation/1_introduction.tex
Rev `.gitignore`, rightly commit all images

\blank [big, force]

\Title {\TitleText}
\blank [big]

\Subtitle {\AuthorText}
\blank [big]

\Subsubtitle {\InstitutionText}

\page [yes]
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Organization}
{
\I Problem Setting

\I Proposed Method

\I Error Analysis

\I Numerical Experiments

\I Conclusion
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Background}
{
\I Multiple-input multiple-output (MIMO) systems \\ have been suggested in 5G communication.

\I Estimation of MIMO channels results in high complexity, due to a large number of antennae.

\I Meanwhile, millimeter wave (mm-Wave) channels are often sparse due its poor scattering nature

\I According to compressive sensing theory, we might exploit the sparsity to reduce complexity of estimation procedure.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Compressive Sensing}
{
\blank [big]
\externalfigure [compressive-sensing.png] [factor=fit]

\I The signal as shown in is not sparse in the basis on the left, but it is in the basis on the right.

\I Thus, the recovery from just a few components are able to approximate it well.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Work on Compressive Channel Sensing}
{
\I Bajwa, Haupt, Sayeed, and Nowak (2010): MIMO, linear invariant system, using Dantzig Selector (DS)

\I Alkhateeb, Ayach, Leus, Heath (2014): mm-wave channel, adaptive training algorithm for given codebook

\I Alkhateeb, Leus, and Heath (2015): single path mm-wave channel, all-phase-shifter beamformers,
using Orthogonal Matching Pursuit (OMP)

\I Lee, Lee, and Yong (2016): MIMO mm-wave channel, hybrid beamforming, using Orthogonal Matching Pursuit (OMP)
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Our Contribution}
{
\I Consider MIMO mm-wave channel, with hybrid beamforming

\I Use a modified DS (convex programming), instead of OMP (greedy algorithm)

\I Illustrate by experiments that DS has better performance

\I Give an analysis of error norm and successful probability in terms of system parameters
}