### 2020-10-21 presentation/2_setting.tex

\Frame {Channel Model and Beamformers}
{
\I Given channel \m {\M {H} \in \NC \MB {C} ^{N_H \D N_H}} with uniform linear array

\I Complex standard normal noise \m {\m {Z} \in \NC \MB {C} ^{N_H \D N_H}}

\I Tx: Digital precoder \m {\M {F}_B \in \NC \MB {C} ^{N_R \D N_Y}}
\I Tx: Analog precoder \m {\NC \M {F}_R \in \NC \MB {C} ^{N_H \D N_R}}

\I Rx: Digital combiner \m {\M {W}_B \in \NC \MB {C} ^{N_R \D N_H}}
\I Rx: Analog combiner \m {\M {W}_R \in \NC \MB {C} ^{N_Y \D N_R}}

\I Magnitude constraint \m {\Nm {\M {F}_R \SB {n_H, n_R}} = 1}, \m {\Nm {\M {W}_R \SB {n_R, n_H}} = 1} 
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {System Diagram}
{
\blank [big]
\externalfigure [system.png] [factor=fit]

\I Problem (a): Recover \m {\M {H}} with knowledge of \\
\m {\M {Y} :=\M {W}_B \M {W}_R \RB {\M {H} \M {F}_R \M {F}_B +\M {Z}} }?

\I Problem (b): Design \m {\M {W}_R}, \m {\M {W}_B}, \m {\M {F}_R}, and \m {\M {F}_B}?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Cast to Linear System}
{
\I Introduce the DFT matrix \m {\M {K} \in \NC \MB {C} ^{N_H \D N_H}}, and write
\Disp {
\NC \V {y}
:= \NC \Rm {vec} \SB {\M {Y}},
\Q \M {P}
:= \RB {\M {F}_B^\Tr \M {F}_R^\Tr \m {K}^\ast} \otimes \RB {\M {W}_B \M {W}_R \M {K}}, \NR
\NC \V {g}
:= \NC \Rm {vec} \SB {\M {K}^\Adj \M {H} \M {K}},
\Q \V {z}
:= \Rm {vec} \SB {\M {Z}} \NR
}

\I Then we have a underdetermined linear system:
\Disp {
\NC \V {y}
=\NC \M {P} \V {g} +\V {z} \NR
}

\I When the signal is sparse, few measurements may be sufficient for the reconstruction
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Terminology}
{
\I \m {\V {g}} is called \Emph {\m {s}-sparse}, if at most \m {s} of its components are nonzero.

\I We say that \m {\M {P}} satisfies the \Emph {restricted isometry property} (RIP) of sparsity \m {s} with respect to \m {0 < \d_s < 1}, if, for all \m {s}-sparse \m {\V {g}},
\Disp {
\NC \RB {1-\d_s} \VNm {\V {g}} _2^2
\leq \NC \VNm {\m {P} \V {g}} _2^2
\leq \RB {1+\d_s} \VNm {\V {g}} _2^2
}

\I We may think \m {\M {P}} to be almost unitary up to relative error \m {\d_s}.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Applying Dantzig Selector}
{
\I We try to apply Dantzig Selector to our problem:

\I Input \m {\M {P} \in \MB {C} ^{N_Y^2 \D N_H^2}}, \m {\V {y} \in \MB {C} ^{N_H^2}}, \m {\g_{\Ss {DS}} \geq 0}.

\I Calculate
\Disp {
\NC \Hat {\V {g}}
\LA \NC \startcases
\NC \Min {\V {g}'}  \MC \VNm {\V {g}'} _1 \NR
\NC \Rm {subject} \; \Rm {to} \Q \MC \VNm {\M {P}^\Adj \RB {\V {y} -\M {P} \V {g}'}} _\infty \leq \g_{\Ss {DS}} \NR
\stopcases \NR
}

\I Output \m {\Hat {\M {H}} \LA \M {K} \Rm {vec}^{-1} \SB {\Hat {g}} \M {K}^\Adj}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Design of Beamformers}
{
\I Let us set each entry of \m {\m {F}_B} (resp.\ \m {\m {W}_B}) to be i.i.d.\ complex standard normal, up to a design constant

\I And set each entry of \m {\m {F}_R} (resp.\ \m {\m {W}_R}) be uniformly distributed in argument and unity magnitude, up to a design constant

\I We will argue that \m {\M {F} _R ^\Tr}, \m {\M {F} _B ^\Tr}, \m {\M {W} _R}, and \m {\M {W} _B} all have RIP, thus \m {\M {P}} has too.

\I Fix sparsity \m {s = L \RB {\log N_H}^2} for concreteness
}