### 2020-10-21 presentation/3_proof.tex

\Frame {Error Analysis}
{
\I What's the difference of true value \m {\V {h}} and estimated value \m {\Hat {\V {h}}}?

\I Should account for the fact that \m {\V {g}} is not strictly sparse

\I Should justify that all four beamformers, \m {\M {F}_B, \M {F}_R, \M {W}_B, \M {W}_R}, have RIP for high probability, thus \m {\m {P}} does too.

\I Should investigate the Big \m {\MC {O}} of error norm by bounding failure events mentioned above.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {The Bound}
{
\I In conclusion, it holds for probability \m {p} that
\Disp {
\NC \VNm {\Hat {\V {g}} - \V {g}} _2
%
\leq \NC 8 \R {L} \RB {\log N_H} ^{3/2} \NR
}
%
\I For design values
\Disp {
\NC N_Y \gtrsim \NC 4 \log N_H, \Q
N_R \gtrsim 16 L \RB {\log N_H}^2 \NR
}
%
\I And it holds that
\Disp {
\NC 1 - p
\lesssim \NC 2 N_H ^{-1/3} \NR
}
%
\I Unfortunately, the values are too big to simulate.
}