### 2020-11-03 treatise/3_2_confirming_restricted_isometry.tex

\startsection [title={Restricted isometry of effective beamformer}]

\startsubsection [title={Design of beamformers}]

Our plan is setting the four beamformers to be i.i.d.\ random matrix, hoping that the resulting \m {P} has RIP.
Let us make it simple and set each entry of \m {\M {F} _B} to be i.i.d.\ normal r.v.\ with mean 0, standard deviation \m {1/2}, multiplied by a normalizing constant \m {\l_B >0}.
We know that the magnitude \m {\RB {\M {F} _B} _{\SB {n_R, n_Y}}} follows Rayleigh distribution.
\m {\RB {\M {W} _B ^\Tr} _{\SB {n_Y, n_R}}} is set in an entirely analolous manner.
In addition, let \m {\RB {\M {F} _R} _{\SB {n_H, n_R}}} be uniformly distributed on the unit circle on the complex plane.
\m {\RB {\M {W} _R ^\Tr} _{\SB {n_Y, n_R}}} is set in an entirely analolous manner.

It turns out we have results that guarantee that both \m {\M {F} _B ^\Tr} and \m {\M {F} _R ^\Tr} are RIP, and same can be said of \m {\M {W} _B} and \m {\M {W} _R}.
Consider a sparse \m {\V {u}} or unity norm, then we are sure that \m {\M {F} _R ^\Tr \V {u}} has almost unity norm.
If, furthermore, \m {\M {F} _R ^\Tr \V {u}} is sparse, we see that \m {\M {F} _B ^\Tr \M {F} _R ^\Tr \V {u}} has almost unity norm, and the RIP of \m {\M{P}} also follows by construction.

However, it is not clear that \m {\M {F} _B ^\Tr \M {F} _R ^\Tr}, the product of two RIP matrices, have RIP, nor that, \m {\M {P}}, the result of Kronecker product, is RIP.
In fact, the statement seems to be false in general.
But all is not lost.
Can we argue instead that \m {\M {F} _B ^\Tr \M {F} _R ^\Tr \V {u}} has similar norm with \m {\M {F} _R ^\Tr \V {u}}?

\stopsubsection

\startsubsection [title={Operator norm of digital beamformers}]

Use \m {\chi _{i,j}} to denote complex standard normal r.vs., and to simplify notation, let them be independent if they have different indices.
With this,
%
\DispNum {C':C':WB:j2} {
\NC \VNm {\M {W} _B \V {u}} _2
=\NC \R {
   \sum _{i=0} ^{N_B-1} \VNm {\sum _{j=0} ^{N_R-1} \F {1} {\R{N_B}} \chi _{i,j} \V {u} _{j}} _2 ^2} \NR
}
%
Essentially, we seek for a large- deviation result for i.i.d.\ sum of chi-square r.vs.\ now.
To so so, we may apply Lemma 1 in Laurent and Massart \cite [LaM00] for equal weights.
By some calculation, we have upper tail inequality, for \m {x > 0},
%
\DispNum {P:x:ex:ex} {
\NC \MB {P} \SB {\VNm {\M {W} _B \V {u}} _2 ^2 \geq 1 + 2 \R {\F {x} {N_B}} + \F {2x} {N_B}}
\leq \NC \Ss {e} ^{-x} \NR
}
%
And lower tail inequality
%
\DispNum {P:x:ex:ex!} {
\NC \MB {P} \SB {\VNm {\M {W} _B \V {u}} _2 ^2 \leq 1 - 2 \R {\F {x} {N_B}}}
\leq \NC \Ss {e} ^{-x} \NR
}
%
With substitution \m {\d_s =16x^2 /N_B^2}, the upper tail can be simplified with \m {x <N_B}, and we get the following bound for failure probability.

\Result
{Proposition}
{
For \m {0 <\d_s <1},
%
\DispNum {W:1:ds:ds} {
\NC \Nm {\VNm {\M {W} _B \V {u}} _2 ^2 - 1}
\geq \NC \d_s \NR
}
holds for probability \m {p}, with
\DispNum {1:p:2e:s4} {
\NC 1 -p
\leq \NC 2 \Ss {e} ^{-N_B \R {\d_s} /4} \NR
}
}

\stopsubsection

\startsubsection [title={Operator norm of analog beamformers}]

Here, we consult Haviv and Regev, “The restricted isometry property of subsampled Fourier matrices” collected in \cite [KlM17].
We need their Thm.\ 4.5; it gives a RIP guarantee w.r.t.\ \m {\d_s}, and probability lower bound, for DFT submatrix.
The estimation they give involve is expressed in terms of big- \m {\O} and big- \m {O}, and since our methodology is 
However, it can be seen that their argument allows precise constants, and they just did not spell out, which we briefly explain.
Roughly, they use Chernoff-Hoeffding bounds, but write the exponent in the form of \m {\Omega}.
Starting in their Thm.\ 2.2, we may substitute (say) \m {C \leftarrow 1/3} in place of their big- \m {\Omega}.
The conclusions, their Thm.\ 3.7 and Thm.\ 4.5, may be changed accordingly.
If modified with respect to our setting, the condition
%
\DispNum {N:R:sd':NH} {
\NC N_R
\geq \NC \F {s} {\d_s^2} \RB {\log \F{1}{\d_s}}^2 \RB {\log \F{s}{\d_s}}^2 \log N_H \NR
}
%
guarantees a failure probability \m {\d_s /N_H s}.
We would like to drop \m {\log \SB {1 /\d_s}^2} and the denominator \m {\d_s} in \m {\log \SB {s /\d_s}^2} in \Rf {N:R:sd':NH}.

\Result
{Theorem}
{
Suppose
%
\DispNum {N:R:sd'NH} {
\NC N_R
\geq \NC \F {s} {\d_s^2} \RB {\log s}^2 \log N_H \NR
}
%
Then \m {\M {F}_R} (resp.\ \m {\M {W} _R}) has \m {\d_s}-RIP for probability \m {p}, with
\DispNum {1:p:ex:d's} {
\NC 1 -p
\leq \NC \RB {\F {\d_s} {N_H s}} ^{1/3} \NR
}
}

\stopsubsection

\startsubsection [title={Confirming the restricted isometry}]

So, how do we relate RIP of \m {\M {F} _R ^\Tr}, \m {\M {F} _B ^\Tr}, \m {\M {W} _R}, and \m {\M {W} _B} to that of \m {\M {P}}?
Notice, by the nature of operator norm,
\DispNum {P:2:FB:WR} {
\NC \VNm {P} _2
=\NC \VNm {\M {F}_B^\Tr \M {F}_B^\Tr} _2 \VNm {\M {W}_B \M {W}_R} _2 \NR
}
Indeed, if it were not for the sparsity constraint, if both \m {\M {F}_B^\Tr \M {F}_B^\Tr} and \m {\M {F}_B^\Tr \M {F}_B^\Tr} are \m {\d_s}-RIP, \m {\M {P}} has approximately \m {2\d_s}-RIP.
However, it is doubtful to assume that the Kronecker product of two RIP matrices exhibit RIP.

We suggest the following argument.
If all angles \m {\th_l, \f_l} are on exactly on the grid, then, with overwhelming probability, it is possible to write
\DispNum {G:G:Ud:dU} {
\NC \M {G}
=\NC \M {U} \Rm {diag} \SB {\V {d}'} \M {U} ^\Adj \NR
}
for some \m {S}-sparse \m {\V {d}'} and unitary \m {\M {U}}.
Clearly,
\DispNum {G:2:g2:g2} {
\NC \VNm {\M {G}} _2
=\NC \VNm {\V {g}} _2 \NR
%
\NC =\NC \VNm {\Rm {abs} \SB {\V {g}}} _2 \NR
}
where \m {\Rm {abs}} denotes elementwise taking absolute value.
Then if we notice that \m {\M {W} _B \M {W} _R \M {K} \M {U}} and \m {\M {U} ^\Adj \M {K}^\Adj \M {F} _R^\Adj \M {F} _B^\Adj} has \m {\d_s}-RIP too, we see
\DispNum {W:2:WB:d's} {
\NC \NC \VNm {\M {W} _B \M {W} _R \M {H} \M {F} _R \M {F} _B} _2 \NR
%
\NC =\NC \VNm {
   \M {W} _B \M {W} _R \M {K} \M {U} \D
   \Rm {diag} \SB {\Rm {abs} \SB {\V {d}'}} \D
   \M {U} ^\Adj \M {K} ^\Adj \M {F} _R \M {F} _B} _2 \NR
%
\NC =\NC \VNm {
   \M {W} _B \M {W} _R \M {K} \M {U}
   \R {\Rm {abs} \SB {\V {d}'}} \D
   \R {\Rm {abs} \SB {\V {d}'}}
   \M {U} ^\Adj \M {K} ^\Adj \M {F} _R \M {F} _B} _2 \NR
%
\NC =\NC \VNm {
   \M {W} _B \M {W} _R \M {K} \M {U}
   \R {\Rm {abs} \SB {\V {d}'}}} _2 \D
   \VNm {\M {F} _B ^\Adj \M {F} _R ^\Adj \M {K} \M {U}
   \R {\Rm {abs} \SB {\V {d}'}}} _2 \NR
%
\NC \eqsim \NC \VNm {\M {G}} _2 \RB {1 +2\d_s} \NR
}
Thus, even we cannot make sure that \m {\M {P}} is \m {\d_s}-RIP with respect to all possible \m {\V {g}}'s, at least \m {\M {P}} has RIP with respect to relevant \m {\V {g}} in accordance to the form of \Rf {H:H:l0:ql}.

\stopsubsection

\stopsection