### 2020-11-03 treatise/acknowledgement.tex

\starttitle [title={Acknowledgement}]

For various reasons, I took very long time writing this treatise.
In the start, I came across the field of compressive sensing in 2018 when I was the TA of the course of Convex Optimization, and also when I was an intern in MediaTek.
I told Prof.\ Su of the proposal, which by then was hardly more than a vague stream of conciousness.
I thank him for always being generous to let students find their own topic of interest, and for having encouraged me to continue writing what would become of the present treatise, even not without procrastination and interruption.

I also want to thank Chu-Tong Liu, The PhD student of Prof.\ Su's, and also an aquaintence of mine since the freshman year, in that he gave me various suggestions when I was stuck and was tolerant when I did not, or even fail at all, to deliver a proper presentation.

Furthermore, I thank my parents, who in recent years was supportive of whatever goal I set up to do (or at least claim to work on), if slightly too credulous regarding my ability.
It was a tremendous luck that I did not have to mind the financial matters: something I ought to cherish, but did so only in later years.
I was happy to have been free to keep a sense of curiosity of the intricacy of the universe.

November 4, 2020

\stoptitle