### 2020-11-08 presentation/2_setting.tex

% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {System Configuration}
{
\blank [big]
\externalfigure [system.png] [factor=fit]

\I Estimation of Multiple-input multiple-output (MIMO) systems results in high complexity, due to a large number of antennae.

\I Problem (a): How to recover the channel with fewer measurements?

\I Problem (b): How to design precoders and combiners?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Channel Model and Beamformers}
{
\I Given channel \m {\M {H} \in \NC \MB {C} ^{N_H \D N_H}} with uniform linear array

\I Complex standard normal noise \m {\m {Z} \in \NC \MB {C} ^{N_H \D N_H}}

\I Tx: Digital precoder \m {\M {F}_B \in \NC \MB {C} ^{N_R \D N_Y}}
\I Tx: Analog precoder \m {\NC \M {F}_R \in \NC \MB {C} ^{N_H \D N_R}}

\I Rx: Digital combiner \m {\M {W}_B \in \NC \MB {C} ^{N_R \D N_H}}
\I Rx: Analog combiner \m {\M {W}_R \in \NC \MB {C} ^{N_Y \D N_R}}

\I Magnitude constraint \m {\Nm {\M {F}_R \SB {n_H, n_R}} = 1}, \m {\Nm {\M {W}_R \SB {n_R, n_H}} = 1} 
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Problem Description}
{
\I We have \m {\M {Y} =\M {W}_B \M {W}_R \RB {\M {H} \M {F}_R \M {F}_B +\M {Z}}}
%
\I Problem (a): How to recover \m {\M {H}} with knowledge of \m {\M {Y}}?

\I Can we measure less than \m {N_H} times, but recover successfully for most cases?

\I Problem (b): How to design \m {\M {W}_R}, \m {\M {W}_B}, \m {\M {F}_R}, and \m {\M {F}_B}?

\I What kind of matrices faciliates our goal?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Cast to Linear System}
{
\I Introduce the DFT matrix \m {\M {K} \in \NC \MB {C} ^{N_H \D N_H}}, and write
\Disp {
\NC \V {y}
:= \NC \Rm {vec} \SB {\M {Y}},
\Q \M {P}
:= \RB {\M {F}_B^\Tr \M {F}_R^\Tr \m {K}^\ast} \otimes \RB {\M {W}_B \M {W}_R \M {K}}, \NR
\NC \V {g}
:= \NC \Rm {vec} \SB {\M {K}^\Adj \M {H} \M {K}},
\Q \V {z}
:= \Rm {vec} \SB {\M {Z}} \NR
}

\I We have an underdetermined system \m {\V {y} =\M {P} \V {g} +\V {z}}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Compressive Sensing}
{
\blank [big]
\externalfigure [compressive-sensing.png] [factor=fit]

\I If \m {\V {g}} is sparse, few measurements may suffice for reconstruction.

\I Can we find a basis on which \m {\V {g}} is sparse?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Literature on Compressive Channel Sensing}
{
\I Bajwa, Haupt, Sayeed, and Nowak (2010): MIMO, linear invariant system, using Dantzig Selector (DS)

\I Alkhateeb, Ayach, Leus, Heath (2014): mm-wave, adaptive training algorithm for given codebook

\I Alkhateeb, Leus, and Heath (2015): single path mm-wave, all-phase-shifter beamformers, using Orthogonal Matching Pursuit (OMP)

\I Lee, Lee, and Yong (2016): MIMO mm-wave, hybrid beamforming, using Orthogonal Matching Pursuit (OMP)
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Question}
{
\I Candès and Tao (2007) proposed a performance bound for DS, then can DS do better than OMP?

\I Does convex programming (DS) trade accuracy with greedy algorithms (OMP)?

\I May we verify a revised performance bound in the concrete case?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Our Contribution}
{
\I Consider MIMO mm-wave channel, with hybrid beamforming

\I Use a modified DS (convex programming), instead of OMP (greedy algorithm)

\I Illustrate by experiments that DS has better performance

\I Give an analysis of error norm and successful probability in terms of system parameters
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Design of Beamformers}
{
\I We set each entry of \m {\m {F}_B} (resp.\ \m {\m {W}_B}) to be i.i.d.\ complex standard normal, up to a design constant

\I We set each entry of \m {\m {F}_R} (resp.\ \m {\m {W}_R}) be uniformly distributed in argument and unity magnitude, up to a design constant

\I We argue that \m {\M {F} _R ^\Tr}, \m {\M {F} _B ^\Tr}, \m {\M {W} _R}, and \m {\M {W} _B} all have RIP, thus \m {\M {P}} has too.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Applying Dantzig Selector}
{
\I Fix sparsity \m {s = L \RB {\log N_H}^2} for definiteness.

\I Input \m {\M {P} \in \MB {C} ^{N_Y^2 \D N_H^2}}, \m {\V {y} \in \MB {C} ^{N_H^2}}, \m {\g_{\Ss {DS}} \geq 0}.

\I Calculate
\Disp {
\NC \Hat {\V {g}}
\LA \NC \startcases
\NC \Min {\V {g}'}  \MC \VNm {\V {g}'} _1 \NR
\NC \Rm {subject} \; \Rm {to} \Q \MC \VNm {\M {P}^\Adj \RB {\V {y} -\M {P} \V {g}'}} _\infty \leq \g_{\Ss {DS}} \NR
\stopcases \NR
}

\I Output \m {\Hat {\M {H}} \LA \M {K} \Rm {vec}^{-1} \SB {\Hat {g}} \M {K}^\Adj}
}