### 2020-11-08 presentation/5_conclusion.tex

\Frame {Conclusion}
{
\I We estimate MIMO mm-Wave channel \m {\M {H}} by using a modified DS.

\I We justify DS in present scenario with almost-sparse \m {\V {g}}, and show the RIP of effective beamformer \m {\M {P}}.

\I We propose a error norm analysis of DS in terms of parameters \m {N_H}, \m {N_Y}, and \m{L}, and investigate the asymptotic successful probability.

\I DS may be converted to be a linear program to reduce complexity.

\I While DS does appear superior, it has complexity much greater than OMP.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {References (1/2)}
{
{\tfx
\I A Alkhateeb, O E Ayach, G Leus, R W Heath, \Quote {Channel Estimation and Hybrid Precoding for Millimeter Wave Cellular Systems}, \It {IEEE Journal of Selected Topics in Signal Processing} 8(5), 2014

\I A Alkhateeb, G Leus, and R W Heath, \Quote {Compressed sensing based multi-user millimeter wave systems: How many measurements are needed?}, in \It {2015 IEEE International Conference on Acoustics, Speech and Signal Processing}.

\I W U Bajwa, J Haupt, A M Sayeed, and R Nowak, \Quote {Compressed channel sensing: A new approach to estimating sparse multipath channels}, \It {Proceedings of the IEEE}, 98(6), 2010

\I S Boyd and L Vandenberghe, \It {Convex optimization}, Cambridge U Press, 2004.
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {References (2/2)}
{
{\tfx
\I E Candès and J Romberg, \Quote {\m {\ell_1}-magic: Recovery of Sparse Signals via Convex Programming}, retrieved online, 2005.

\I E Candès and T Tao, \Quote {The Dantzig Selector: Statistical estimation when \It {p} is much larger than \It {n}}, \It {The Annals of Statistics} 35(6), 2007.

\I M P Friedlander and M A Saunders, \Quote {Discussion: the Dantzig selector: statistical estimation when \It {p} is much larger than \It {n}}, \It {The Annals of Statistics}, 35(6), 2007.

\I J Lee, G-T Gil, and Y H Lee, \Quote {Channel estimation via orthogonal matching pursuit for hybrid MIMO systems in millimeter wave communications}, \It {IEEE Transactions on Communications}, 64(6), 2016.

\I J A Tropp and A C Gilbert, \Quote {Signal recovery from random measurements via orthogonal matching pursuit}, \It {IEEE Transactions on information theory}, 53(12), 2007.
}
}