### 2020-11-08 treatise/acknowledgement.tex

\starttitle [title={Acknowledgement}]

For various reasons, I took very long time writing this treatise.
In the beginning, I came across the field of compressive sensing in 2018 when I was the TA of the course of Convex Optimization, and also when I was an intern in MediaTek.
I told Prof.\ Su of the proposal, which by then was hardly more than a vague stream of conciousness.
I thank him for always letting students find their own topic of interest, and for having encouraged me to continue writing what would become of this treatise, even not without procrastination and interruption.
In addition to the stipend he kindly supplies, the warm faux-wood floor, comfortable sofas, and ergonomic chairs in the lab also helped.

The present treatise is far from being remarkable, but I suppose the course of devising, clarifying, speculating, verifying, programming, debugging, googling, and polishing a work does teach some resilience to a previously weak mind.
The three and half years are punctuated by Bachi martial art practice session, teaching assistant's preparation and scoring, chess club meeting, MediaTek internship, and the two semester courses of Real Analysis and Advanced Statistical Inference, which are recollections I too hold dearly.

I also thank Chu-Tong Liu, The PhD student of Prof.\ Su's, who I acquainted since in the freshman year, in that he provided me various suggestions when I was stuck and was lenient when I did not, or even fail at all, to deliver a proper presentation.
And I enjoy the meals, trips, and board games with the fellow lab students, whose discussion, concern, jokes, and sarcasms are particularly enlightening when one is undergoing a writer's block.

After all, I thank my parents, who in recent years was supportive of whatever goal I set up to do (or at least claim to work on), if slightly too credulous regarding my ability.
It was a tremendous luck that I did not have to mind the financial matters for now, which is something I ought to treasure, but did so only in later years.
I was happy to have been free to keep a sense of curiosity of the intricacy of the universe.

November 4, 2020

\stoptitle