### 2020-11-09 presentation/4_conclusion.tex

\Frame {Conclusion}
{
\I We estimate MIMO mm-Wave channel \m {\M {H}} by using modifed DS.

\I We justify DS under relaxed sparsity of \m {\V {g}}, and demonstrate the RIP of effective beamformer \m {\M {P}}.

\I We analyze the error norm of DS in terms of \m {N_H}, \m {N_Y}, and \m{L}, and approimate the asymptotic failure probability.

\I By simulation we see that DS can be cast into be a linear program to reduce complexity, and appear to outperform other methods.

\I DS trades off low time and space complexity high precision, and for OMP the opposite is true.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {References (1/2)}
{
{\tfx
\I A Alkhateeb, O E Ayach, G Leus, R W Heath, \Quote {Channel Estimation and Hybrid Precoding for Millimeter Wave Cellular Systems}, \It {IEEE Journal of Selected Topics in Signal Processing} 8(5), 2014

\I A Alkhateeb, G Leus, and R W Heath, \Quote {Compressed sensing based multi-user millimeter wave systems: How many measurements are needed?}, in \It {2015 IEEE International Conference on Acoustics, Speech and Signal Processing}.

\I W U Bajwa, J Haupt, A M Sayeed, and R Nowak, \Quote {Compressed channel sensing: A new approach to estimating sparse multipath channels}, \It {Proceedings of the IEEE}, 98(6), 2010

\I S Boyd and L Vandenberghe, \It {Convex optimization}, Cambridge U Press, 2004.
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {References (2/2)}
{
{\tfx
\I E Candès and J Romberg, \Quote {\m {\ell_1}-magic: Recovery of Sparse Signals via Convex Programming}, retrieved online, 2005.

\I E Candès and T Tao, \Quote {The Dantzig Selector: Statistical estimation when \It {p} is much larger than \It {n}}, \It {The Annals of Statistics} 35(6), 2007.

\I M P Friedlander and M A Saunders, \Quote {Discussion: the Dantzig selector: statistical estimation when \It {p} is much larger than \It {n}}, \It {The Annals of Statistics}, 35(6), 2007.

\I J Lee, G-T Gil, and Y H Lee, \Quote {Channel estimation via orthogonal matching pursuit for hybrid MIMO systems in millimeter wave communications}, \It {IEEE Transactions on Communications}, 64(6), 2016.

\I J A Tropp and A C Gilbert, \Quote {Signal recovery from random measurements via orthogonal matching pursuit}, \It {IEEE Transactions on information theory}, 53(12), 2007.
}
}