### 2020-11-10 presentation/2_analysis.tex

\Frame {Error Analysis}
{
\I Can we bound \m {\VNm {\V {h} -\Hat {\V {h}}} _2}?

\I Does it matter that \m {\V {g}} is not strictly sparse?

\I If \m {\M {F}_B, \M {F}_R, \M {W}_B, \M {W}_R}, all have restricted isometry, does \m {\m {P}} have too?

\I Big \m {\MC {O}} probability of failure events?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {The Bound}
{
\I We argue in the treatise that it holds for probability \m {p} that
\Disp {
\NC \VNm {\Hat {\V {g}} - \V {g}} _2
%
\eqsim \NC \MC {O} \SB {\R {L} \RB {\log N_H} ^{3/2}} \NR
}
%
\I For ideal values
\Disp {
\NC N_B \gtrsim \NC 4 \log N_H, \Q
N_R \gtrsim 16 L \RB {\log N_H}^2 \NR
}
%
\I And it holds that
\Disp {
\NC 1 - p
\eqsim \NC \MC {O} \SB {\Ss {e} ^{-N_B /12}}
 \NR
}
%
\I Unfortunately, the values are too big to simulate.
}