### 2020-11-10 presentation/4_conclusion.tex

\Frame {Conclusion}
{
\I We estimate MIMO mm-Wave channel \m {\M {H}} by using modified DS.

\I We justify DS under relaxed condition on sparsity of \m {\V {g}}, and demonstrate the RIP of effective beamformer \m {\M {P}}.

\I We analyze the error norm of DS in terms of \m {N_H}, \m {N_B}, and \m{L}, and approximate the asymptotic failure probability.

\I By simulation we see that DS appears to be superior to other methods.

\I DS trades off low time and space complexity for high precision.
For OMP, the opposite is true.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {References (1/2)}
{
{\tfx
\I A Alkhateeb, O E Ayach, G Leus, R W Heath, \Quote {Channel Estimation and Hybrid Precoding for Millimeter Wave Cellular Systems}, \It {IEEE Journal of Selected Topics in Signal Processing} 8(5), 2014

\I A Alkhateeb, G Leus, and R W Heath, \Quote {Compressed sensing based multi-user millimeter wave systems: How many measurements are needed?}, in \It {2015 IEEE International Conference on Acoustics, Speech and Signal Processing}.

\I W U Bajwa, J Haupt, A M Sayeed, and R Nowak, \Quote {Compressed channel sensing: A new approach to estimating sparse multipath channels}, \It {Proceedings of the IEEE}, 98(6), 2010

\I S Boyd and L Vandenberghe, \It {Convex optimization}, Cambridge U Press, 2004.
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {References (2/2)}
{
{\tfx
\I E Candès and J Romberg, \Quote {\m {\ell_1}-magic: Recovery of Sparse Signals via Convex Programming}, retrieved online, 2005.

\I E Candès and T Tao, \Quote {The Dantzig Selector: Statistical estimation when \It {p} is much larger than \It {n}}, \It {The Annals of Statistics} 35(6), 2007.

\I M P Friedlander and M A Saunders, \Quote {Discussion: the Dantzig selector: statistical estimation when \It {p} is much larger than \It {n}}, \It {The Annals of Statistics}, 35(6), 2007.

\I J Lee, G-T Gil, and Y H Lee, \Quote {Channel estimation via orthogonal matching pursuit for hybrid MIMO systems in millimeter wave communications}, \It {IEEE Transactions on Communications}, 64(6), 2016.

\I J A Tropp and A C Gilbert, \Quote {Signal recovery from random measurements via orthogonal matching pursuit}, \It {IEEE Transactions on information theory}, 53(12), 2007.
}
}