### 2020-11-10 treatise/4_simulation.tex

\startchapter [title={Simulation}]

In this chapter we shall compare \Rf {d:2:OL:H3} with numerical experiments.
For the convex optimization, we choose the Python library CVXPY.
For the implementation details, see Diamond and Boyd \cite [DiB16]
A direct implementation of Algorithm [7] leads to extraordinarily large complexity, but Candès and Romberg \cite [CaR05] showed that DS can be cast as a linear program, which we illustrate below.

\startsection [title={Method}]

\startsubsection [title={Some Notation}]

For brevity we shall represent complex vector and matrices by real ones, consisting of its real and imaginary parts.
For \m {\V {x} \in \MB {C} ^{M}}, define the real representation \m {\MS {R} \SB {\V {x}}} of \m {\V {x}} to be
%
\DispNum {R:x:R2:M1} {
\NC \MS {R} \SB {\V {x}}
\in \NC \MB {R} ^{2M} \NR
\NC \MS {R} \SB {\V {x}} _{\SB {m}}
= \NC \startcases
\NC \MF {Re} \SB {\V {x} _{\SB {m'}}}, \MC m =2m' \NR
\NC \MF {Im} \SB {\V {x} _{\SB {m'}}}, \MC m =2m'+1 \NR
\stopcases \NR
\NC m' 
= \NC  0, 1, 2, \ldots, M-1 \NR
}
%
The injection is obvious, and we may define \m {\MS {R} ^{-1}} so that
%
\DispNum {R:1:x::x:} {
\NC \MS {R} ^{-1} \SB {\MS {R} \SB {\V {x}}}
=\NC \V {x} \NR
}
%
Accordingly, the following generalization to complex matrices is valid, once we call the ring representation of complex numbers.
For \m {\M {A} \in \MB {C} ^{M_1 \D M_2}}, define real representation \m {\MS {R} \SB {\M {A}}} of \m {\M {A}} to be
%
\DispNum {R:A:R2:21} {
\NC \MS {R} \SB {\M {A}}
\in \NC \MB {R} ^{2M_1 \D 2M_2} \NR
\NC \MS {R} \SB {\M {A}} _{\SB {m_1,m_2}} =
\NC \startcases
\NC \MF {Re} \SB {\M {A} _{\SB {m_1',m_2'}}}, \MC (m_1, m_2) = (2m_1', 2m_2') \NR
\NC \MF {Im} \SB {\M {A} _{\SB {m_1',m_2'}}}, \MC (m_1, m_2) = (2m_1'+1, 2m_2') \NR
\NC -\MF {Im} \SB {\M {A} _{\SB {m_1',m_2'}}}, \MC (m_1, m_2) = (2m_1', 2m_2'+1) \NR
\NC \MF {Re} \SB {\M {A} _{\SB {m_1',m_2'}}}, \MC (m_1, m_2) = (2m_1'+1, 2m_2'+1) \NR
\stopcases \NR
\NC m_1 
= \NC 0, 1, 2, \ldots, M_1 -1 \NR
\NC m_2 
= \NC 0, 1, 2, \ldots, M_2 -1 \NR
}

With these, we define
%

\DispNum {y:y:Ry:Ny} {
\NC \T {\V {y}}
= \NC \MS {R} \SB {\V {y}}
\in \MB {R} ^{2 N_B^2} \NR
%
\NC \T {\V {g}}
= \NC \MS {R} \SB {\V {g}}
\in \MB {R} ^{2 N_H^2} \NR
%
\NC \T {\M {P}}
= \NC \MS {R} \SB {\M {P}}
\in \MB {R} ^{2 N_B^2 \D 2 N_H^2} \NR
%
\NC \T {\M {P}} ^\Adj
= \NC \MS {R} \SB {\M {P} ^\Adj}
\in \MB {R} ^{2 N_B^2 \D 2 N_H^2} \NR
%
\NC \T {\V {z}}
= \NC \MS {R} \SB {\V {z}}
\in \MB {R} ^{2 N_B^2} \NR
}
%
so, by construction,
%
\DispNum {y:y:Pg:gz} {
\NC \V {\T {y}}
= \NC \M {\T {P}} \V {\T {g}} +\V {\T {z}} \NR
}
%
And denote \m {\V {1}} to be the all-\m{1} vector, and \m {\V {0}} the all-\m{0} vector, whose dimension will be inferred from context.

\stopsubsection

\startsubsection [title={A Linear Program}]

We can now see that DS is equivalent to a linear program.

\Result
{Algorithm}
{
\startitemize[n]
\item Input \m{\M {P} \in \MB {C} ^{ N_B^2 \D N_H^2}}, \m{\V {y} \in \MB {C} ^{ N_B^2}}, \m {\g_{\Rm {DS}} > 0}.
%
\item Define \m {\T {\V {y}}, \T {\V {g}}, \T {\M {P}}, \T {\M {P}} ^\Adj, \T {\V {z}}}, according to \Rf {y:y:Ry:Ny}.
%
\item Calculate
\DispNum {g:f:gf:S1} {
\NC \Hat {\T {\V {g}}}, \Hat {\T {\V {f}}}
\LA \NC \startcases
   \NC \Min {\T {\V {g}}', \T {\V {f}}'}  \MC \IP {\V {1}, \T {\V {f}}'} \NR
   \NC \Rm {subject} \; \Rm {to} \MC \T {\V {g}}' \preceq \T {\V {f}}' \NR
   \NC \MC - \T {\V {g}}' \preceq \T {\V {f}}' \NR
   \NC \MC \T {\M {P}}^\Adj \T {\M {P}} \T {\V {g}}' \preceq \T {\M {P}}^\Adj \T {\V {y}} + \g_{\Ss {DS}} \V {1} \NR
   \NC \MC - \T {\M {P}}^\Adj \T {\M {P}} \T {\V {g}}' \preceq - \T {\M {P}}^\Adj \T {\V {y}} + \g_{\Ss {DS}} \V {1} \NR
\stopcases \NR
}
\item Convert
%
\DispNum {g:::R1:1g} {
\NC \Hat{\V {g}}
\LA \NC \MS {R} ^{-1} \SB {\Hat{\T{\V {g}}}} \NR
}
\item Calculate
%
\DispNum {G:G:ve:1g} {
\NC \Hat {G}
\LA \NC \Rm {vec}^{-1} \SB {\Hat {g}} \NR
}
\item Calculate
%
\DispNum {H:H:KG:GK} {
\NC \Hat {\M {H}}
\LA \NC \M {K} \Hat {\M {G}} \M {K}^\Adj \NR
}
\item Output \m {\Hat {\M {H}}}.
\stopitemize
}

\stopsubsection

\startsubsection [title={A Two-Stage Version}]

We also tried another version of DS, in which DS is applied twice in estimated nonzero components.
The two-stage method was briefly mentioned \cite [CaT07] but without further justification by numerical experiments.
To apply the method, we first apply DS as before.
Then, we extract largest components of the estimated vector, and apply DS again.
Furthermore, we extract largest components of the second estimated vector, and apply Moore–Penrose inverse.
The final indices set has to be scrambled back corresponding to the original ones.

\Result
{Algorithm}
{
\startitemize[n]
\item Let \m {\g_{\Rm {DS}} \geq 0} be given, and \m {\M {P}}

\item Set
\Disp {
\NC N_0
=\NC 2 N_H^2, \NR
\NC N_2
=\NC \lfloor 4 \log N_H \rfloor, \NR
\NC N_1
=\NC \lfloor \R {N_0 N_2} \rfloor \NR
}

\item Apply DS to \m {\V {y}, \M {P}} to get \m {\Hat {\V {g}}_0}, and call the \m {N_1} largest component of \m {\Hat {\V {g}}_0} to be \m {\V {g}_1}, and corresponding columns of \m {\M {P}} to be \m {\M {P} _1}.

\item Apply DS to \m {\V {y}, \M {P}_1} to get \m {\Hat {\V {g}}_1}, and call the \m {N_2} largest component of \m {\Hat {\V {g}}_1} to be \m {\V {g}_2}, and corresponding columns of \m {\M {P}} to be \m {\M {P} _2}.

\item Apply Moore–Penrose inverse to \m {\V {y}, \M {P} _2} to get \m {\Hat {\V {g}}_2}, which corresponds to \m {\Hat {\V {g}}}.
\stopitemize
}

Unfortunately, the two-stage method does not always appear to be better than the one-stage method, and we did not include the method in the plots.
It might have something to do with appropriately scaling \m {\g_{\Rm {DS}}} with respect to \m {N_H},
and we do not know what most reasonable choice of \m {N_1, N_1, N_2} is.

\stopsubsection

\startsection [title={Result}]

\startsubsection [title={Settings}]

Beside DS, we have simulated OMP for three different stop conditions, Lasso, and Moore Penrose pseudoinverse (abbreviated LS for least square).
For the plot of assorted methods, \m {8} values of \m {\s} are considered; for the plots of only OMP, \m {9} values are considered; for the plots of only OMP, \m {6} values are considered.

To get some idea of the order of magnitude of noise-to-signal level, plug in some actual numbers.
Consider only path loss in the simplest form according to Friis Law.
Suppose the power of mobile phone antenna is 0.25 W,
the carrier frequency is 5GHz,
the base station is \m {1.5} km away,
the noise is \m {–40} dB W.
If so, the noise-to-signal ratio is
\DispNum {1:8:02:24} {
\NC 10^{-4} \F {1} {0.25} \R {\F {5 \D 10^9 \D 4 \pi \D 1500} {3 \D 10^8}}
=\NC 0.224 \NR
}
In the following simulation, we use dimensionless noise-to-signal levels starting with \m {2^{-4}}, and being multiplied with powers of \m {2}.

We take \m {N_H} to be \m {8, 12, 16, 20, 24}, respectively.
For assorted plots, two series of plots are simulated.
One series for \m {N_B = \lfloor N_H /2 \rfloor}, giving \m {4, 6, 8, 10, 12};
one series for \m {N_B = \lfloor N_H /3 \rfloor}, giving \m {3, 4, 5, 7, 8}.
For OMP and DS only plots, we only consider \m {N_B = \lfloor N_H /3 \rfloor}.
On the other hand, \m {N_R} always set to be \m {\lfloor \RB {\log N_H}^2 \rfloor}.
Unfortunately, this is very far from achieving the design values \Rf {N:B:4l:H2}, and this may be part of the reason the result is not as successful as expected.

Other parameters are fixed in these experiments; 
The number of grid of quantization of analog precoder is \m {16}.
The number of paths \m {L} is \m {4}.
The wavelength of carrier \m {\l _{\Rm {ant}}} is set to be \m {0.1}.
and the antenna spacing \m {d _{\Rm {ant}}} is \m {0.2}.
Only their ratio matters.

For parameters related to the convex optimization algorithm, the tolerance of absolute error in the primal-dual algorithm used in CVXPY is set as \m {10^{-4}}, and the tolerance of relative error, as \m {10^{-3}}.
The maximal number of iteration of CVXPY is set to be \m {32}.
On the other hand, the maximal number of iteration of OMP is set to be \m {4 N_B}.

Denote the threshold used in DS program to be \m {\g_{\Rm {DS}}}, and similar threshold of Lasso to be \m {\g_{\Rm {Lasso}}}.
Different values are produced by multiplying powers of \m {2}.

In the plots of assorted methods, we set \m {\g_{\Rm {DS}} = 2 \R {\log N_H}} for concreteness, as suggested in \cite [CaT07].
For sake of comparison, \m {\g_{\Rm {Lasso}} = 2 \R {\log N_H}} too.
For OMP, suggested values in Cai and Wang \cite [CaW11] is illuminating, where they (in their Theorem 7) considered \m {\ell _2}-norm, and (in their Theorem 8), \m {\ell _\infty}-norm.
We simply take \m {\h_{\Rm {OMP}} = 2 \R {\log N_H}} for \m {\infty}-norm constraint, \m {\h_{\Rm {OMP}} = \R {3 N_B}} for 2-norm constraint.

In the plots of DS only or OMP only, we vary the values of \m {\g_{\Rm {DS}}} and \m {\h_{\Rm {OMP}}} to observe the effects of thresholds.

Each data point for DS and Lasso is repeated for \m {256} times, and taken average.
Other methods are repeated for more times: OMP for \m {4 \D 256} times, LS for \m {12 \D 256} times.

For performance metric, we follow Lee, Gil, and Lee \cite [LGL16],
\DispNum {h:h:lo:vg} {
\NC \T {\chi}
=\NC \RB {
   \F {\log_2 {\VNm {\V {h} -\Hat {\V {h}}} _2}}
   {\log_2 {\VNm {\V {h}}_2}}
} _{\Ss {avg}}, \NR
}
However, we note that when \m {\VNm {\V {h}}_2} is small, this can blow up.
We do not know, either, whether \m {\T {\chi}} is a good indicator of channel capacity, since we do not fix a channel model in our investigation.

The parameters related to precision of the Newton step may be adjusted from CVXPY's class methods.
We set the maximum absolute tolerance to be \m {5 \D 10 ^ {-7}},
the maximum absolute tolerance to be \m {5 \D 10 ^ {-6}},
the maximum feasible tolerance to be \m {5 \D 10 ^ {-7}}.
It remains to be tested what values are most suitable for our purposes.
If the tolerance parameters are set too small, the program often gives overflowing values, perhaps because it was not able to find feasible solutions.
On the other hand, if they are set too big, the Newton steps are so imprecise that the program fails too.

\stopsubsection

\startsubsection [title={Assorted Plots}]

In the following, we plot the reciprocal of noise level vs relative error norm, for \m {N_H = 8, 12, 16, 20, 24}, respectively.
Figures 3 to 6 have \m {N_H = 3 N_B}, and figures 7 to 10 has \m {N_H = 2 N_B}.

\blank [big]
\externalfigure [assorted-wide-small-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {Assorted methods, \m {N_H = 3 N_B, N_H = 12}, error.}
\blank [big]
%
\blank [big]
\externalfigure [assorted-wide-medium-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {Assorted methods, \m {N_H = 3 N_B, N_H = 16}, error.}
%
\blank [big]
\externalfigure [assorted-wide-big-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {Assorted methods, \m {N_H = 3 N_B, N_H = 20}, error.}
%
\blank [big]
\externalfigure [assorted-wide-very-big-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {Assorted methods, \m {N_H = 3 N_B, N_H = 24}, error.}
%
\blank [big]
\externalfigure [assorted-narrow-small-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {Assorted methods, \m {N_H = 2 N_B, N_H = 12}, error.}
\blank [big]
%
\blank [big]
\externalfigure [assorted-narrow-medium-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {Assorted methods, \m {N_H = 2 N_B, N_H = 16}, error.}
\blank [big]
%
\blank [big]
\externalfigure [assorted-narrow-big-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {Assorted methods, \m {N_H = 2 N_B, N_H = 20}, error.}
\blank [big]
%
\blank [big]
\externalfigure [assorted-narrow-very-big-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {Assorted methods, \m {N_H = 2 N_B, N_H = 24}, error.}
\blank [big]

As for the plot of time taken in minutes, we give only one example of \m {N_H = 16}.
%
\blank [big]
\externalfigure [assorted-medium-time.png] [wfactor=fit, hfactor=fit]
\FigureCaption {Assorted methods, \m {N_H = 3 N_B, N_H = 16}, time.}
\blank [big]

We give one additional plot for extreme low signal level, in order to observe the trend of curves.
\blank [big]
\externalfigure [assorted-medium-low-signal-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {Assorted methods, \m {N_H = 3 N_B, N_H = 16}, low signal level, error.}
\blank [big]

\stopsubsection

\startsubsection [title={Plots of DS only}]

Again, we plot the reciprocal of noise level vs relative error norm, for \m {N_H = 8, 12, 16, 20, 24}, respectively, and \m {N_H = 3 N_B} for each of them.
Here, \m {\T {\chi}} is shown too.
%
\blank [big]
\externalfigure [ddss-small-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {DS, \m {N_H = 3 N_B, N_H = 12}, error.}
\blank [big]
%
\blank [big]
\externalfigure [ddss-medium-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {DS, \m {N_H = 3 N_B, N_H = 16}, error.}
\blank [big]
%
\blank [big]
\externalfigure [ddss-big-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {DS, \m {N_H = 3 N_B, N_H = 20}, error.}
\blank [big]
%
\blank [big]
\externalfigure [ddss-very-big-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {DS, \m {N_H = 3 N_B, N_H = 24}, error.}
\blank [big]
%
As for the plot of time taken in minutes, we give only one example of \m {N_H = 16}.
%
\blank [big]
\externalfigure [ddss-medium-time.png] [wfactor=fit, hfactor=fit]
\FigureCaption {DS, \m {N_H = 3 N_B, N_H = 16}, time.}
\blank [big]

\stopsubsection

\startsubsection [title={Plots of OMP only}]

Again, we plot the reciprocal of noise level vs relative error norm, for \m {N_H = 8, 12, 16, 20, 24}, respectively, and \m {N_H = 3 N_B} for each of them.

%
\blank [big]
\externalfigure [oommpp-small-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {OMP, \m {N_H = 3 N_B, N_H = 12}, error.}
\blank [big]
%
\blank [big]
\externalfigure [oommpp-medium-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {OMP, \m {N_H = 3 N_B, N_H = 16}, error.}
\blank [big]
%
\blank [big]
\externalfigure [oommpp-big-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {OMP, \m {N_H = 3 N_B, N_H = 20}, error.}
\blank [big]
%
\blank [big]
\externalfigure [oommpp-very-big-error.png] [wfactor=fit, hfactor=fit]
\FigureCaption {OMP, \m {N_H = 3 N_B, N_H = 24}, error.}
\blank [big]
%
As for the plot of time taken in minutes, we give only one example of \m {N_H = 16}.
%
\blank [big]
\externalfigure [oommpp-medium-time.png] [wfactor=fit, hfactor=fit]
\FigureCaption {OMP, \m {N_H = 3 N_B, N_H = 16}, time.}
\blank [big]

\stopsubsection

\startsubsection [title={Discussion}]

In the assorted plots, DS steadily outperforms other methods.
From lowest to highest error, it usually goes like this: DS, LS, Lasso, OMP.
Though the result is encouraging for our purpose, we wonder why OMP is so bad.
Did we make mistakes?
We do not think so.
A crucial reason might be that in Lee, Gil, and Lee \cite [LGL16], the angles of departure and arrival is distributed within a cluster of standard deviation 15 degrees, and in ours, they are uniform on the unit circle.
One is then tempted to say, we suggested a low complexity beamformer and high complexity estimation method, while they suggested a high complexity beamformer and low complexity estimation method, and the low complexity solution does not seem effective in our settings.
It remains to see, however, how OMP and DS compare, if done exactly under their settings.

However, it is not clear how higher \m {\s} (dimensionless noise level) corresponds to lower \m {\T {\chi}} (simulated error).
It seems when \m {\s} is further increased (leftmost of the plots), the error might also increase rapidly, but more parameters have to be simulated for a conclusion.

Varying thresholds of DS (namely \m {\g_{\Rm {DS}}}) and OMP (namely \m {\h_{\Rm {OMP}}}) appears not to affect the estimation significantly.
Nevertheless, it seems \m {\g_{\Rm {DS}}} is usually best chosen in the order of magnitude of \m {\s}, which is reasonable.
Sometimes the iteration stabilizes much earlier than is constrained by maximum iteration number.

Sadly, we remark that CVXPY sometimes gives overflowing outputs, some of them as large as \m {10^{11}}.
This indicates perhaps some typical-looking output may in fact be unreliable.
We thus discard outputs larger than a given threshold, say \m {10^4}, and simply return the answer to be a Moore–Penrose inverse.
It is unknown whether there are better methods.

We are not sure either, unfortunately, why \m {\T {\chi}} does not agree \m {\chi}, and why \m {\chi} does not appear to increase significantly as \m {\s} increases.
As shown above, \m {\chi} is almost constant around \m {1}, thus \m {\T {\chi}} is an overestimation when \m {\s} is high and underestimation when \m {\s} is low.
A possible explanation is that the non-sparsity of \m {\V {g}} undermines the analysis in chapter 3, despite our attempts to account for that effect.
These matters remain to be investigated in the future.

We include Figure 12 for completeness's sake, where extreme low signal level is considered.
Starting with \m {-30} dB, error of all methods grow significantly.
However, the data may not be entirely reliable, since as we said we automatically discard overflowing values when the program blows up.
All in all, the trend remains to be reproduced perhaps by simulating on a better device.

We end this section by showing two cases where DS works or fails, in figures 23 and 24.
In both figures the true values of \m {\M {H}} entries are ordered by magnitude, and corresponding estimated values in the first and second stage are plotted against them.
When DS fails, the soft thresholding did not rule out smaller entries, thus a second application of DS on the wrong index set failed too.
When DS succeeded, the soft thresholding gave a reasonable guess of the nonzero components, thus a second application of DS on this index set refined the estimated values.

\blank [big]
\externalfigure [scatter-ddss-failure.png] [wfactor=300]
\FigureCaption {A scatter plot of position versus magnitude for a case in which DS failed.}
\blank [big]

\blank [big]
\externalfigure [scatter-ddss-success.png] [wfactor=300]
\FigureCaption {A scatter plot of position versus magnitude for a case in which DS succeeded.}
\blank [big]

\stopsection

\startsection [title={Complexity}]

\startsubsection [title={Asymtotic Analysis}]

We want to know the complexity of DS and OMP.
In general, complexity of convex programs is difficult to analyze, in that it is not easily to determine how many iterations were done.
Boyd and Vandenberghe \cite [BoV04] analyzed complexity of inequality constrained Newton method, in which self-concordance is assumed (p.496ff, also p.531).
In our case, Candès and Romberg \cite [CaR05] pointed out that the complex DS can be cast into an second order cone programming (SOCP), and where self concordance holds.
Diamond and Boyd (writers of CVXPY) \cite [DiB16] explained that convex programs are usually converted into primal-dual problems.
The library we called in the simulation is ECOS, where Newton method is used.
Of course, we are not certain there is no method faster than Newton method, and implementation details of CVXPY may change the conclusion too.

We attemp to give a heuristic argument regarding the complexity of Newton steps.
We say the function being minimized is \m {\VNm {\V {g}}_1}.
Let \m {\V {g} _0} denote the starting value of \m {\V {g}}, and \m {\V {g} ^{\star}} the the point of convergence.
Then Boyd and Vandenberghe \cite [BoV04] (p.505, Eqn.9.56) gave a bound of the number of steps.
If we just take that as the complexity bound for DS, it becomes
%
\DispNum {C:S:C0:1e'} {
\NC C_{\Rm {DS}}
= \NC \RB {C_0 \VNm {\V {g}_0 -\V {g} ^{\star}}_1
+ \log_2 \log_2 \F {1} {\e}} C_{\Rm {step}} \NR
}
%
Here \m {C_0} is constant related to implementation of Newton method, and \m {\e} the tolerance of error, and \m {C_{\Rm {step}}} is the complexity of Newton step.
According to their analysis (Eqn.9.57), for most cases \m {-\log_2 \log_2 \e} can be bounded by \m {6}.
Also, \m {C_0} usually assumes the value of about several hundred.
There was an example they gave in which \m {C_0 =375}, where, with their notation,
\m {C_0 =(20 -8\a) / (\a \b \RB {1 -2\a}^2}), where \m {\a = 0.1, \b = 0.8, \e =0.01}, and \m {\a, \b} are parameters used in the Backtrack Line Tracing algorithm).

Every Newton step involves a inversion of \m {\MS {O} \SB {\M {P} ^\Adj \M {P}}}, having the same order of magnitude to the time needed for matrix multiplication.
Since the dimension of \m {\M {P} ^\Adj \M {P}} is \m {N_H^2}, the cost of inversion may be roughly considered to be \m {\MS {O} \SB {N_H^4}}, as we know from newer results on the complexity of matrix inversion.
That is, we suppose \m {C_{\Rm {step}} = N_H^4}.

We take some liberty to assume that, at the start the iteration, the Moore–Penrose inverse was taken to be the initial value of \m {\V {g}_0}, namely
%
\DispNum {g:0:gL:Py} {
\NC \V {g}_0
=\NC \V {g}_{\Rm {LS}} \NR
\NC = \NC \RB {\M {P} ^\Adj \M {P}} ^{-1} \M {P} ^\Adj \V {y} \NR
}
%
By above, we may write the complexity of DS to be
%
\DispNum {C:S:Og:g1} {
\NC C_{\Rm {DS}}
=\NC \MS {O} \SB {\VNm {\V {g}_{\Rm {LS}} -\V {g} ^\star} _1} N_H^4 \NR
}
%
Also assume
%
\DispNum {g:g:g;:;g} {
\NC \V {g} ^\star
\approx \NC \V {g} \NR
}
%
Indeed, this is the main thesis of our investigation.
And by restricted isometry, \m {\M {P}} has unity-normed, almost orthogonal columns, so
%
\DispNum {P:P:IN:H2} {
\NC \M {P} ^\Adj \M {P}
\approx \NC I _{N_{H}^2} \NR
}
%
Thus, by \Rf {g:0:gL:Py}, \Rf {C:S:Og:g1}, and \Rf {g:g:g;:;g},
%
\DispNum {C:S:Og:z1} {
\NC C_{\Rm {DS}}
=\NC \MS {O} \SB {
   \VNm {\V {g} +\RB {\M {P} ^\Adj \M {P}} ^{-1} \M {P} ^\Adj \V {z}
   -\V {g}} _1} N_H^4 \NR
\NC =\NC \MS {O} \SB {\VNm {\M {P} ^\Adj \V {z}} _1} N_H^4 \NR
}

We see that \m {\RB {\M {P} ^\Adj \V {z}} _{\SB {j}}} observes complex standard normal distribution.
If we agree that, for high probability,
%
\DispNum {O:j:sO:O1} {
\NC \MS {O} \SB {\VNm {\M {P} ^\Adj \V {z}} _{\SB {j}}}
= \NC \s \MS {O} \RB {1} \NR
}
Then we simply have
%
\DispNum {C:S:ON:H6} {
\NC C_{\Rm {DS}}
=\NC \MS {O} \SB {N_{H}^6} \NR
}

On the other hand, Tropp and Gilbert \cite [TrG07a] discussed the complexity of OMP in the companion paper \cite [TrG07b].
They gave
%
\DispNum {C:P:ON:NY} {
\NC C_{\Rm {OMP}}
=\NC \MS {O} \SB {N_H^2 \log N_B} \NR
}
%
Here \m {\log N_B} is small in our experiments, and if we dropped some constants and \m {\log N_B}, then we may assume \m {\MS {O} \SB {N_H^2}}.

Lastly, since Lasso involves the same target function as DS, the same argument is valid, giving \m {C_{\Rm {Lasso}} =\MS {O} \SB {N_H^6}}.

\stopsubsection

\startsubsection [title={Runtime Statistics}]

All plots included in this treatise, in total, took probably one month to run.
Figures 6 and 10 each took about \m {3} days.
The computer has \m {32} GB RAM, and Intel Core i9 processor.
Simulation ran on Linux subsystem 2 of Windows 10.

Use \m {t_{\Rm {OMP}} \SB {N_H}}, \m {t_{\Rm {Lasso}} \SB {N_H}}, \m {t_{\Rm {DS}} \SB {N_H}} to denote their respective time taken.
We reproduce the arithmetic average (including different \m {\s}) of simulated values below, in seconds.
\DispNum {t:2:95:55} {
\NC t_{\Rm {OMP}} \SB {12} \NC = 0.0119,
\FourQ t_{\Rm {OMP}} \SB {16} = 0.0297, \NR
\NC t_{\Rm {OMP}} \SB {20} \NC = 0.0600,
\FourQ t_{\Rm {OMP}} \SB {24} = 0.155, \NR
%
\NC t_{\Rm {Lasso}} \SB {12} \NC = 0.0436,
\FourQ t_{\Rm {Lasso}} \SB {16} = 0.176, \NR
\NC t_{\Rm {Lasso}} \SB {20} \NC = 0.471,
\FourQ t_{\Rm {Lasso}} \SB {24} = 1.26, \NR
%
\NC t_{\Rm {DS}} \SB {12} \NC = 0.357,
\FourQ t_{\Rm {DS}} \SB {16} = 2.07, \NR
\NC t_{\Rm {DS}} \SB {20} \NC = 29.8,
\FourQ t_{\Rm {DS}} \SB {24} = 91.5, \NR
}
Simulation indeed shows that \m {t_{\Rm {DS}}} grows tremendously with (in our case) \m {N_H}, agreeing the criticism by Friedlander and Saunders \cite [FrS07] on the high complexity of DS.
It is not surprising, regarding the fact that the convex program has to find the Newton step by calculating (in our case) \m {\M {P}}, a matrix of dimension \m {\M {N}_H^2}.
The matrix inversion is probably the bottleneck of complexity.
However, as we noted above, that different tolerance level of and maximal iteration number may result in vastly different computation time.

Meanwhile, \m {t_{\Rm {Lasso}}} and \m {t_{\Rm {OMP}}} are often an order of magnitude lower than \m {t_{\Rm {DS}}}.
In general, \m {t_{\Rm {DS}} \gg t_{\Rm {Lasso}} \gg t_{\Rm {OMP}}}.
It is now wonder that, given that OMP is a greedy algorithm, the time OMP takes is negligibly small.
But it is also notable that, while Lasso has form similar to DS, it takes much lower time.

Assuming the complexity of each of three methods is polynomial time, we calculate a least square fit for the power of complexity, giving the following estimated values, on which we add a tilde.

\DispNum {t:P:NH:83} {
\NC \T {t}_{\Rm {OMP}} \NC \eqsim \MS {O} \SB {N_H^{3.6}} \NR
\NC \T {t}_{\Rm {Lasso}} \NC \eqsim \MS {O} \SB {N_H^{4.8}} \NR
\NC \T {t}_{\Rm {DS}} \NC \eqsim \MS {O} \SB {N_H^{8.3}} \NR
}

\stopsubsection

\stopsection

\stopchapter