### 2020-11-10 treatise/conclusion.tex

\startchapter [title={Conclusion}]

This treatise investigates effective estimation of the millimeter wave, multiple-input, multiple-output (mm-wave MIMO) channel.
We adopt the Dantzig Selector (DS) rather than Orthogonal Matching Pursuit (OMP), which is most commonly used in literature, by exploiting the sparsity of the problem in spacial Fourier domain.
We bounded the expected error norm following the spirit of Candès and Tao \cite [CaT07].
To do so, we evaluate the sparsity, confirm the restricted isometry of sensing matrix \m {\M {P}}, and generalize the method for complex vectors.
Our bound, on a general setting, is characterize in terms of physical parameters, such as the number of path \m {L} and channel dimension \m {N_H}.
Simulation is done, and DS's performance is compared to OMP, Lasso, and LS.
DS outperforms others, at the cost of higher complexity, agreeing our analysis.
Put concisely, DS trades off time and space complexity for precision, and we hope that our inquiry provides a welcome alternative to the present literature.

\stopchapter