### 2020-12-31 treatise-pdflatex/abstract.tex

\starttitle [title={Abstract}]

Multiple-input-multiple-output (MIMO) is the future of wireless communication, but the large number of antennae makes it difficult to estimate its channel state information.
Since millimeter wave (mm-Wave) band, which is often used together with MIMO, is known to exhibit sparsity, results on compressive sensing suggest that, on mm-Wave band, fewer channel measurements shall suffice.
Current literature prefers orthogonal matching pursuit (OMP) due to its lower complexity.
It is interesting to note that Dantzig Selector (DS), claimed to have various performance guarantee, has seldom been examined.

In this treatise, we consider a mm-Wave MIMO channel, and a hybrid beamforming structure at both the transmitter and receiver end.
We demonstrate that DS can be applied on a entrywise i.i.d.\ random beamformers, and give a bound for its expected error norm, which holds for high probability.
DS is cast as a linear program, and simulation is done, showing that, in our problem setting, DS is superior in terms of relative error norm.

\stoptitle