### 2020-12-31 treatise-pdflatex/foreword.tex

\starttitle [title={Foreword}]

I came across the field of compressive sensing in 2018, to begin with, when I was the TA of the course of Convex Optimization, and read about it again as an intern in MediaTek.
Prof.\ Su always had students find their own topic of interest, and I thank him for having encouraged me to continue writing what would become of this treatise, even if when I first told him of the proposal, it was hardly more than a stream of conciousness.
In addition to the stipend he kindly supplied, the warm faux-wood floor, comfortable sofas, and ergonomic chairs in the lab also helped.
For various reasons, I took very long time writing this treatise, and I thank the advisor for his tolerance on my procrastination, if that is not too late.

The present treatise is far from being remarkable, but I suppose the course of devising, clarifying, speculating, verifying, programming, debugging, googling, and polishing does teach some resilience to a weak mind.
The three and half years are punctuated by Bāchí martial art practice session, teaching assistant's preparation and scoring, chess club meeting, MediaTek internship, and the two semester courses of Real Analysis and Advanced Statistical Inference, which are recollections I too hold dearly.

In addition, I thank Chu-Tong Liu, a fellow PhD student of Prof.\ Su's, who I have acquainted in the undergraduate freshman year, in that he gave me valuable suggestions when I was stuck, and was lenient when I did not, or even failed at all to, deliver a proper presentation.
I too cherish the decent meals, trips, and board games enjoyed together with fellow lab students, whose discussion, concern, joke, and sarcasm are particularly enlightening when one is undergoing a writer's block.

After all, I thank my parents, who in recent years was supportive of whatever goal I set up to do (or at least claimed to work on), if slightly too credulous in my ability.
It so was a tremendous luck that I did not have to mind financial matters in the master student years, which is something I ought to treasure.
I am happy to have kept a sense of curiosity for the intricate universe.

November 4, 2020

\stoptitle