### 2020-12-31 treatise-pdflatex/notation.tex

\starttitle [title={Notation}]

\starttable[|l|l|l|] \HL
Expression \VL Meaning or Definition \VL More Explanation \SR \HL
\m {\Nm {a}} \VL magnitude \VL for real or complex numbers \AR \HL
\m {\mathfrak {Re} \SB {a}} \VL real part \VL for complex numbers \AR \HL
\m {\mathfrak {Im} \SB {a}} \VL imaginary part \VL for complex numbers \AR \HL
\m {\V {a}} \VL vector \VL \AR \HL
\m {\M {A}} \VL matrix \VL \AR \HL
\m {\V {1}} \VL all-1 vector \VL dimension will be specified \AR \HL
\m {\V {0}} \VL all-0 vector \VL dimension will be specified \AR \HL
\m {\V {a} _{\SB {n}}} \VL \m {n}-th component of \m {\V {a}} \VL \AR \HL
\m {\M {A} _{\SB {m,n}}} \VL \m {m,n}-th entry of \m {\M {A}} \VL \AR \HL
\m {\V {a} _{\SB {0}}} \VL the first component of \m {\V {a}} \VL  \AR \HL
\m {\M {A} _{\SB {0,0}}} \VL the upper-left entry of \m {\V {A}} \VL  \AR \HL
\m {\mathbb {K}} \VL either \m {\mathbb {R}} or \m {\mathbb {C}} \VL  \AR \HL
\m {\mathbb {K} ^{N}} \VL Hilbert space \m {\mathbb {K}^N} over \m {\mathbb {K}} \VL  \AR \HL
\m {\mathbb {K} ^{M \D N}} \VL collection of \m {M} by \m {N} matrices \VL  \AR \HL
\m {m\; \mathrm {Mod}\; n} \VL remainder of \m {m} divided by \m {n} \VL with \m {m, n \geq 0} and \m {0 \leq m\; \mathrm {Mod}\; n <n} \AR \HL
\m {\VNm {\V {a}} _p} \VL \m {\ell_p}-norm of \m {\V {a}} \VL \m {p >0} \AR \HL
\m {\VNm {\V {a}} _\infty} \VL the maximum-magnitude \VL  \AR
 \VL component of \m {\V {a}} \VL  \AR \HL
\m {\mathbb {P} \SB {A}} \VL probability of event \m {A} \VL  \AR \HL
\m {\mathbb {E} \SB {X}} \VL expectation of random variable \m {X} \VL  \AR \HL
\m {\mathsf {Var} \SB {X}} \VL variance of random variable \m {X} \VL  \AR \HL
\stoptable