### 2021-02-09 presentation/1_setting.tex
Discuss complexity

\blank [big, force]

\Title {\TitleText}
\blank [big]

\Subtitle {\AuthorText}
\blank [big]

\Subsubtitle {\InstitutionText}
\Subsubtitle {November 2020}

\page [yes]
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {System Configuration}
{
\blank [big]
\externalfigure [system.png] [factor=fit]

\I Estimation of Multiple-input multiple-output (MIMO) systems results in high complexity, due to a large number of antennae.

\I Problem (a): How to reconstruct the channel with fewer measurements?

\I Problem (b): How to generate precoders and combiners?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Channel model and beamforming matrices}
{
\I Given MIMO channel representation \m {\M {H} \in \NC \MB {C} ^{N_H \D N_H}} with uniform linear array, having \m {L} paths, with hybrid beamforming.

\I Complex standard normal noise \m {\M {Z} \in \NC \MB {C} ^{N_H \D N_H}}

\I Tx: Digital precoder \m {\M {F}_B \in \NC \MB {C} ^{N_R \D N_B}}
\I Tx: Analog precoder \m {\NC \M {F}_R \in \NC \MB {C} ^{N_H \D N_R}}

\I Rx: Digital combiner \m {\M {W}_B \in \NC \MB {C} ^{N_R \D N_H}}
\I Rx: Analog combiner \m {\M {W}_R \in \NC \MB {C} ^{N_B \D N_R}}

\I Magnitude constraint \m {\Nm {\M {F}_{R,\SB {n_H, n_R}}} = 1}, \m {\Nm {\M {W}_{R,\SB {n_H, n_R}}} = 1}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Problem Description}
{
\I We have \m {\M {Y} =\M {W}_B \M {W}_R \RB {\M {H} \M {F}_R \M {F}_B +\M {Z}}}
%
\I Problem (a): How to recover \m {\M {H}} with knowledge of \m {\M {Y}}?

\I Can we measure less than \m {N_H^2} times, but recover successfully for most cases?

\I Problem (b): How to generate \m {\M {W}_R}, \m {\M {W}_B}, \m {\M {F}_R}, and \m {\M {F}_B}?

\I What kind of matrices facilitates our goal?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Cast to Linear System}
{
\I Introduce the DFT matrix \m {\M {K} \in \NC \MB {C} ^{N_H \D N_H}}, and write
\Disp {
\NC \V {y}
:= \NC \Rm {vec} \SB {\M {Y}},
\Q \M {P}
:= \RB {\M {F}_B^\Tr \M {F}_R^\Tr \m {K}^\ast} \otimes \RB {\M {W}_B \M {W}_R \M {K}}, \NR
\NC \V {g}
:= \NC \Rm {vec} \SB {\M {K}^\Adj \M {H} \M {K}},
\Q \V {z}
:= \Rm {vec} \SB {\M {W}_B \M {W}_R \M {Z}} \NR
}

\I We have an underdetermined system \m {\V {y} =\M {P} \V {g} +\V {z}}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Compressive sensing}
{
\blank [big]
\externalfigure [compressive-sensing.png] [factor=fit]

\I If \m {\V {g}} is sparse, fewer measurements suffice for reconstruction.

\I Can we find a basis on which \m {\V {g}} is likely to be sparse?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Literature review}
{
\I Bajwa, Haupt, Sayeed, and Nowak (2010): MIMO, linear invariant system, using Dantzig Selector (DS)

\I Destino, Juntti, and Nagaraj (2015): MIMO, digital beamforming, using Lasso

\I Vlachos, Alexandropoulos, and Thompson (2019): MIMO, hybrid beamforming, mm wave, using Lasso

\I Alkhateeb, Leus, and Heath (2015): single path mm wave, all phase shifter beamforming, using Orthogonal Matching Pursuit (OMP)

\I Lee, Lee, and Yong (2016): MIMO mm wave, hybrid beamforming, using Orthogonal Matching Pursuit (OMP)
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Contribution}
{
\I Due to the constraint of hybrid beamforming, designs for pure digital beamforming on sensing matrix might not work.

\I We claim that DS is superior in higher noise and fewer samples than OMP.

\I We propose a low complexity random generation of beamforming matrices, and justify that it may serve as the sensing matrix.

\I We give a asymptotic bound on expected error norm.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Design of Beamformers}
{
\I We set each entry of \m {\m {F}_B} (resp.\ \m {\m {W}_B}) to be i.i.d.\ complex standard normal, up to a design constant

\I We set each entry of \m {\m {F}_R} (resp.\ \m {\m {W}_R}) be uniformly distributed in argument and unity magnitude, up to a design constant

\I We argue that \m {\M {F} _R ^\Tr}, \m {\M {F} _B ^\Tr}, \m {\M {W} _R}, and \m {\M {W} _B} all have restricted isometry, thus \m {\M {P}} has too.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Applying Dantzig Selector}
{
\I Input \m {\M {P} \in \MB {C} ^{N_B^2 \D N_H^2}}, \m {\V {y} \in \MB {C} ^{N_H^2}}, \m {\g_{\Ss {DS}} \geq 0}.

\I Calculate
\Disp {
\NC \Hat {\V {g}}
\LA \NC \startcases
\NC \Min {\V {g}'}  \MC \VNm {\V {g}'} _1 \NR
\NC \Rm {subject} \; \Rm {to} \Q \MC \VNm {\M {P}^\Adj \RB {\V {y} -\M {P} \V {g}'}} _\infty \leq \g_{\Ss {DS}} \NR
\stopcases \NR
}

\I Output \m {\Hat {\M {H}} \LA \M {K} \Rm {vec}^{-1} \SB {\Hat {g}} \M {K}^\Adj}
}