### 2021-02-15 treatise-context/5_conclusion.tex
copy corrected system diagram

\startchapter [title={Conclusion}]

The treatise aims to answer the problem of effective estimation of the MIMO millimeter wave channel by exploiting its sparsity.
To do so, we try the Dantzig Selector (DS) for channel estimation.
We exploited the sparsity of spatial frequency domain, and justified the RIP of effective beamformer for the hybrid structure.
We then proved quantitatively that the expected error norm is bounded for overwhelming probability.
Regarding the complexity, we also suggested several ways to reduce it without losing much performance, as justified by simulation.
Simulation is done, and we see that DS indeed outperforms other methods.

We believe that DS is more accurate while having higher complexity.
When the sample is less sufficient and the noise level is higher, sometimes only DS can recover successfully.
Therefore, DS may be used in the scenario that precision is crucial, and in the scenario that the number of RF chain is more limited or fewer stages of estimation is possible.

\stopchapter