### 2021-03-02 treatise-pdflatex/notation.tex

\chapter*{Notation}

\begin{center}
\begin{tabular}{ l | c | r }
\hline
 Expression  &  Meaning or Definition  &  More Explanation  \\ \hline
 \m {\Nm {a}}  &  magnitude  &  for real or complex numbers  \\ \hline
 \m {\mathfrak {Re} \SB {a}}  &  real part  &  for complex numbers  \\ \hline
 \m {\mathfrak {Im} \SB {a}}  &  imaginary part  &  for complex numbers  \\ \hline
 \m {\V {a}}  &  vector  &   \\ \hline
 \m {\M {A}}  &  matrix  &   \\ \hline
 \m {\V {1}}  &  all-1 vector  &  dimension will be specified  \\ \hline
 \m {\V {0}}  &  all-0 vector  &  dimension will be specified  \\ \hline
 \m {\V {a} _{\SB {n}}}  &  \m {n}-th component of \m {\V {a}}  &   \\ \hline
 \m {\M {A} _{\SB {m,n}}}  &  \m {m,n}-th entry of \m {\M {A}}  &   \\ \hline
 \m {\V {a} _{\SB {0}}}  &  the first component of \m {\V {a}}  &    \\ \hline
 \m {\M {A} _{\SB {0,0}}}  &  the upper-left entry of \m {\V {A}}  &    \\ \hline
 \m {\mathbb {K}}  &  either \m {\mathbb {R}} or \m {\mathbb {C}}  &    \\ \hline
 \m {\mathbb {K} ^{N}}  &  Hilbert space \m {\mathbb {K}^N} over \m {\mathbb {K}}  &    \\ \hline
 \m {\mathbb {K} ^{M \D N}}  &  collection of \m {M} by \m {N} matrices  &    \\ \hline
 \m {m\; \mathrm {Mod}\; n}  &  remainder of \m {m} divided by \m {n}  &  with \m {m, n \geq 0} and \m {0 \leq m\; \mathrm {Mod}\; n <n}  \\ \hline
 \m {\VNm {\V {a}} _p}  &  \m {\ell_p}-norm of \m {\V {a}}  &  \m {p >0}  \\ \hline
 \m {\VNm {\V {a}} _\infty}  &  the maximum-magnitude  &   \\ \hline
   &  component of \m {\V {a}}  &    \\ \hline
 \m {\mathbb {P} \SB {A}}  &  probability of event \m {A}  &    \\ \hline
 \m {\mathbb {E} \SB {X}}  &  expectation of random variable \m {X}  &    \\ \hline
 \m {\mathrm {Var} \SB {X}}  &  variance of random variable \m {X}  &    \\ \hline
\end{tabular}
\end{center}