### 2021-03-14 treatise-context/5_conclusion.tex

\startchapter [title={Conclusion}]

The treatise aims to answer the problem of effective estimating a MIMO millimeter-wave channel by exploiting its sparsity.
To do so, we try the Dantzig Selector (DS) for channel estimation.
We exploited the sparsity in the spatial frequency domain, and justified the RIP of effective beamforming matrix for the hybrid structure.
We then proved quantitatively that the expected error norm is bounded for overwhelming probability.
We also suggested several ways to reduce the complexity without losing much performance, as justified by simulation.
Simulation is done, and we see that DS indeed outperforms other methods.

We corroborate the prediction that DS is nearly optimal, and it may extract more information than other methods do when the sampling rate is low.
When the sample is less abundant and the noise level is higher, sometimes only DS can recover successfully.
Therefore, one may use DS in the scenario that precision is crucial, as well as that the number of RF chains is more limited or fewer stages of estimation is possible.

\stopchapter