### 2021-03-15 treatise-context/5_conclusion.tex

\startchapter [title={Conclusion}]

The treatise aims to answer the problem of effective estimating a MIMO mm-wave channel by exploiting its sparsity.
To do so, we apply the Dantzig Selector (DS), by exploiting the sparsity in the spatial frequency domain, and justified the restricted isometry of the effective beamforming matrix.
We then prove quantitatively that the expected error norm is bounded for overwhelming probability.
We also suggest several ways of reducing the complexity without losing much performance.
Simulation is done, and we see that DS indeed outperforms other methods in most datasets.

We corroborate the prediction that DS is nearly optimal, and it may extract more information than other methods do when the sampling rate is low.
When the sample is less abundant and the noise level is higher, sometimes only DS can recover successfully.
Therefore, one may use DS in the scenario that precision is crucial, as well as the case that the number of RF chains is more limited or fewer stages of estimation are possible.

\stopchapter