### 2021-03-20 presentation/1_setting.tex
Fix presentation typos

\blank [big, force]

\Title {\TitleText}
\blank [big]

\Subtitle {\AuthorText}
\blank [big]

\Subsubtitle {\InstitutionText}
\Subsubtitle {November 2020}

\page [yes]
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {System Configuration}
{
\blank [big]
\externalfigure [system.png] [factor=fit]

\I Consider a multiple-input multiple-output (MIMO) system with hybrid structure, so we have fewer RF chains than the antennas

\I Problem (a): Can we estimate the channel with fewer measurements?

\I Problem (b): What precoders and combiners help the estimation?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Channel model and beamforming}
{
\I Given MIMO channel representation \m {\M {H} \in \NC \MB {C} ^{N_H \D N_H}} with uniform linear array, having \m {L} paths, with hybrid beamforming.

\I Complex standard normal noise \m {\M {Z} \in \NC \MB {C} ^{N_H \D N_H}}

\I Tx: Digital precoder \m {\M {F}_B \in \NC \MB {C} ^{N_{B,t} \D N_{B,t}}}
\I Tx: Analog precoder \m {\NC \M {F}_R \in \NC \MB {C} ^{N_H \D N_{B,t}}}

\I Rx: Digital combiner \m {\M {W}_B \in \NC \MB {C} ^{N_{B,r} \D N_H}}
\I Rx: Analog combiner \m {\M {W}_R \in \NC \MB {C} ^{N_{B,r} \D N_{B,r}}}

\I Magnitude constraint \m {\Nm {\M {F}_{R,\SB {n_H, n_{R,t}}}} = 1}, \m {\Nm {\M {W}_{R,\SB {n_H, n_{R,r}}}} = 1}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Cast to Linear System}
{
\I We have \m {\M {Y} =\M {W}_B \M {W}_R \RB {\M {H} \M {F}_R \M {F}_B +\M {Z}}}
\I Introduce the DFT matrix \m {\M {K} \in \NC \MB {C} ^{N_H \D N_H}}, and write
\Disp {
\NC \V {y}
:= \NC \Rm {vec} \SB {\M {Y}},
\Q \M {P}
:= \RB {\M {F}_B^\Tr \M {F}_R^\Tr \m {K}^\ast} \otimes \RB {\M {W}_B \M {W}_R \M {K}}, \NR
\NC \V {g}
:= \NC \Rm {vec} \SB {\M {K}^\Adj \M {H} \M {K}},
\Q \V {z}
:= \Rm {vec} \SB {\M {W}_B \M {W}_R \M {Z}} \NR
}
\I We have an underdetermined system \m {\V {y} =\M {P} \V {g} +\V {z}}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Problem Description}
{
\I Problem (a): How to recover \m {\M {H}} with knowledge of \m {\M {Y}}?

Can we measure less than \m {N_H^2} times, but recover successfully for most cases?

\I Problem (b): How to generate \m {\M {W}_R}, \m {\M {W}_B}, \m {\M {F}_R}, and \m {\M {F}_B}?

What kind of matrices facilitates our goal?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Compressive sensing}
{
\blank [big]
\externalfigure [compressive-sensing.png] [factor=fit]

\I If \m {\V {g}} is sparse, fewer measurements suffice for reconstruction.

\I Can we find a basis on which \m {\V {g}} is likely to be sparse?
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Literature review}
{
\I Bajwa, Haupt, Sayeed, and Nowak (2010): MIMO, no beamforming, linear invariant system, using Dantzig Selector (DS)

\I Destino, Juntti, and Nagaraj (2015): MIMO, digital beamforming, using Lasso

\I Vlachos, Alexandropoulos, and Thompson (2019): MIMO, hybrid beamforming, mm wave, using Lasso

\I Alkhateeb, Leus, and Heath (2015): single path mm wave, analog beamforming, using Orthogonal Matching Pursuit (OMP)

\I Lee, Lee, and Yong (2016): MIMO mm wave, hybrid beamforming, using Orthogonal Matching Pursuit (OMP)
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Contribution}
{
\I Due to the constraint of hybrid beamforming, sensing matrix designs for fully digital beamforming does not directly work.

\I We propose a low complexity, i.i.d.\ random beamforming matrices, and justify that it may serve as the sensing matrix for DS.

\I We claim that DS is superior in higher noise and fewer samples than OMP.
\I We give a asymptotic bound on expected error norm.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Design of Beamformers}
{
\I We set each entry of \m {\m {F}_B} (resp.\ \m {\m {W}_B}) to be i.i.d.\ complex standard normal, up to a design constant

\I We set each entry of \m {\m {F}_R} (resp.\ \m {\m {W}_R}) be uniformly distributed in argument and unity magnitude, up to a design constant

\I We argue that \m {\M {F} _R ^\Tr}, \m {\M {F} _B ^\Tr}, \m {\M {W} _R}, and \m {\M {W} _B} all have restricted isometry, thus \m {\M {P}} has too.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Applying Dantzig Selector}
{
\I Input \m {\M {P} \in \MB {C} ^{N_{B,t} N_{B,r} \D N_H^2}}, \m {\V {y} \in \MB {C} ^{N_H^2}}, \m {\g_{\Ss {DS}} \geq 0}.

\I Calculate
\Disp {
\NC \Hat {\V {g}}
\LA \NC \startcases
\NC \Min {\V {g}'}  \MC \VNm {\V {g}'} _1 \NR
\NC \Rm {subject} \; \Rm {to} \Q \MC \VNm {\M {P}^\Adj \RB {\V {y} -\M {P} \V {g}'}} _\infty \leq \g_{\Ss {DS}} \NR
\stopcases \NR
}

\I Output \m {\Hat {\M {H}} \LA \M {K} \Rm {vec}^{-1} \SB {\Hat {g}} \M {K}^\Adj}
}