### 2021-03-29 treatise-pdflatex/foreword.tex

\chapter* {Foreword}

I came across the field of compressive sensing in 2018, to begin with, when I was the TA of the course on convex optimization by Prof.\ Borching Su, and read about it again as an intern in MediaTek.
Prof.\ Hsuan-Jung Su always had students find their own topic of interest, and I thank him for having encouraged me to continue writing what would become of this treatise, even if when I first told him of the proposal, it was hardly more than a stream of consciousness.
In addition to the stipend he kindly supplied, the warm faux-wood floor, comfortable sofas, and ergonomic chairs in the lab are also impressive.
For various reasons, I took very long time writing this treatise, and I thank the advisor for his tolerance on my procrastination, if that is not too late.
And at my presentation, Prof.\ Borching Su and Prof.\ I-Hsiang Wang, the oral defense committee, gave a lot suggestions, which are equally indispensable.

The present treatise is far from being remarkable, but I suppose the course of devising, clarifying, speculating, verifying, programming, debugging, googling, and polishing does teach some resilience to a weak mind.
The three and half years are punctuated by Bachi martial art practice session, TA preparation and scoring, chess club meeting, MediaTek internship, and the two semester courses of Real Analysis and Advanced Statistical Inference, which are recollections I hold dearly.

In addition, I thank Chu-Tong Liu, a fellow PhD student, who I have acquainted in the undergraduate freshman year, in that he gave me valuable helps when I was stuck, and was lenient when I did not deliver a proper presentation.
I too cherish the decent meals, trips, and board games enjoyed together with lab students, whose discussion, concern, joke, and sarcasm are particularly enlightening when one is undergoing a writer's block.

After all, I thank my parents, who in recent years was supportive of whatever goal I set up to do (or at least claimed to work on), if slightly too credulous in my ability.
It so was a tremendous luck that I did not have to mind financial matters in the master student years.
Therefore, I am happy to have kept a sense of curiosity for the intricate universe.

November 4, 2020