### 2021-04-28 presentation/3_simulation.tex
pdflatex: Shrink figure width

% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Reducing the complexity}
{
\I Casting to a linear program (not used now)

\I Successively estimate the nonzero components (not used now)

\I Use the basis pursuit denoising form

\Disp {
\NC \Hat {\V {g}}
\LA \NC \Min {\V {g}'} 
\CB {
\VNm {\V {g}'} _1 
+ \l_{\Ss {DS}} \VNm {\M {P}^\Adj \RB {\V {y} -\M {P} \V {g}'}} _\infty
} \NR
}
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Error norm (1/6)}
{
\I \m {N_B=2, N_R=3, N_{H,t}=6, N_{H,r}=8}, 6 stages
\blank [big]
\externalfigure [error-small-more-tall-six-usual.png] [width=11cm]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Error norm (2/6)}
{
\I \m {N_B=2, N_R=3, N_{H,t}=8, N_{H,r}=6}, 6 stages
\blank [big]
\externalfigure [error-small-more-wide-six-usual.png] [width=11cm]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Error norm (3/6)}
{
\I \m {N_B=4, N_R=6, N_{H,t}=12, N_{H,r}=16}, 6 stages
\blank [big]
\externalfigure [error-small-more-tall-six-usual.png] [width=11cm]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Error norm (4/6)}
{
\I \m {N_B=4, N_R=6, N_{H,t}=16, N_{H,r}=12}, 6 stages
\blank [big]
\externalfigure [error-small-more-wide-six-usual.png] [width=11cm]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Error norm (5/6)}
{
\I \m {N_B=6, N_R=8, N_{H,t}=18, N_{H,r}=24}, 6 stages
\blank [big]
\externalfigure [error-small-more-tall-six-usual.png] [width=11cm]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Error norm (6/6)}
{
\I \m {N_B=6, N_R=8, N_{H,t}=24, N_{H,r}=18}, 6 stages
\blank [big]
\externalfigure [error-small-more-wide-six-usual.png] [width=11cm]
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Discussion}
{
\I DS outperforms other methods in the high noise case and less measurement

\I The relaxed sparsity may have compromised its effectiveness.

\I DS has complexity much higher than OMP, and somewhat higher than Lasso.

\I Unfortunately, the values are too small to verify the asymptotic bound.
}
% XXX % % XXX % % XXX % % XXX % % XXX % % XXX %
\Frame {Complexity}
{
\I Let \m {C} denote big-O estimation of complexity, then
\Disp {
\NC C_{\Rm {OMP}} =\NC \mathcal {O} \SB {N_H^2 \log N_{B,t} N_{B,r}} \NR
\NC C_{\Rm {Lasso}} =\NC \mathcal {O} \SB {N_H^6} \NR
\NC C_{\Rm {DS}} =\NC \mathcal {O} \SB {N_{H}^6} \NR
}

\I Fitting the exponent of runtime statistics roughly agrees these result.

\I In certain conditions it is pointed out Lasso and DS are equivalent.
}